# Data Processing for SMORES binary datasets for the 2km grid 

Datasets to be processed:

Canyon

Plan: each dataset will have its own code chunk. All processing will live in that code chunk. The top of the document will house functions, data pathways, and other pieces that will be utilised throughout the script. 


```{r, set-up}
#load libraries
library(tidyverse)
library(sf)
#library(nominatimlite)
library(arcgis)
library(arrow)
library(sfarrow)

# 2km base grid
grid.2km <- sf::st_read(
  dsn = "G:\\Shared drives\\NMFS NWC FRAM GIS\\GIS_Base\\Grid.gdb",
  layer = "grid2km_ETsquare"
)

# st_write_parquet(
#   grid.2km,
#   "C:\\GitHub\\SMORES\\data\\2km\\grid_full.parquet"
# )

# get base grid as a dataframe
base_grid_df <- grid.2km %>%
  st_drop_geometry()

# write_parquet(
#   base_grid_df,
#   "C:\\GitHub\\SMORES\\data\\5km\\base_grid_df.parquet"
# )

# set standard coordinate reference system if using 2 km square grid
crsOut <- st_crs(grid.2km)

# calculate the area of the grid cells
grid_2km <- grid.2km %>%
  mutate(cell_area = st_area(.))

# file path where collections of layers lives
melissa_file_path <- "G:\\Shared drives\\NMFS NWC OEI GIS\\ArcGIS\\Projects\\OWEC\\p30\\nccos_share_CA2.gdb"

```

# Canyons

```{r canyons}
# read file in
Canyons <- sf::st_read(
  dsn = melissa_file_path,
  layer = "SubmarineCanyons_WestCoast_diss"
) %>%
  st_transform(crsOut)

Canyons.Score = 0.1

# intersect and calculate the area of the overlapping pieces
Canyons.grid <- sf::st_intersection(Canyons, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
canyon_threshold <- Canyons.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.Canyons = Canyons.Score) %>%
  select(CellID_2km, Score.Canyons)

canyon_scored <- grid.2km %>%
  inner_join(canyon_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.Canyons) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
canyons_scored_long <- pivot_longer(
  canyon_scored,
  cols = starts_with(c("0", "1")),
  names_to = "Canyon",
  values_to = "Score.Canyon"
) %>%
  sf::st_transform(4326) %>%
  select(-Canyon) %>%
  st_zm()

st_write_parquet(
  canyons_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\canyon_scored_2km_new.parquet"
)

canyon_old_2 <- st_read_parquet(
  "C:\\GitHub\\SMORES\\data\\2km\\canyon_scored_full.parquet"
)

plot(canyons_scored_long)
```

# Deep Sea Coral

```{r canyons}
# read file in
DSC.RobustHigh <- sf::st_read(
  dsn = melissa_file_path,
  layer = "DSC_HabitatSuitability_RobustHigh"
) %>%
  st_transform(crsOut)

DSC.RH.Score = 0.1

# intersect and calculate the area of the overlapping pieces
DSC.RobustHigh.grid <- sf::st_intersection(DSC.RobustHigh, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
DSC.RobustHigh_threshold <- DSC.RobustHigh.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.DSC.RH = DSC.RH.Score) %>%
  select(CellID_2km, Score.DSC.RH)

DSC_RH_scored <- grid.2km %>%
  inner_join(DSC.RobustHigh_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.DSC.RH) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
DSC_RH_scored_long <- pivot_longer(
  DSC_RH_scored,
  cols = starts_with(c("0", "1")),
  names_to = "DSC_RH",
  values_to = "Score.DSC_RH"
) %>%
  sf::st_transform(4326) %>%
  select(-DSC_RH) %>%
  st_zm()

st_write_parquet(
  DSC_RH_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\DSC_RH_scored_2km_new.parquet"
)
```


# Seeps

```{r canyons}
# read file in
Seeps <- sf::st_read(
  dsn = melissa_file_path,
  layer = "CascadiaSeeps_wBuffer"
) %>%
  st_transform(crsOut)

Seeps.Score = 0.1

# intersect and calculate the area of the overlapping pieces
Seeps.grid <- sf::st_intersection(Seeps, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
Seeps_threshold <- Seeps.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.Seeps = Seeps.Score) %>%
  select(CellID_2km, Score.Seeps)

Seeps_scored <- grid.2km %>%
  inner_join(Seeps_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.Seeps) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
Seeps_scored_long <- pivot_longer(
  Seeps_scored,
  cols = starts_with(c("0", "1")),
  names_to = "Seeps",
  values_to = "Score.Seeps"
) %>%
  sf::st_transform(4326) %>%
  select(-Seeps) %>%
  st_zm()

st_write_parquet(
  Seeps_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Seeps_scored_2km_new.parquet"
)
```


# Shelf Break

```{r canyons}
# read file in
ShlfBrk <- sf::st_read(
  dsn = melissa_file_path,
  layer = "ContinentalShelfBreak_wBuffer"
) %>%
  st_transform(crsOut)

ShlfBrk.Score = 0.1

# intersect and calculate the area of the overlapping pieces
ShlfBrk.grid <- sf::st_intersection(ShlfBrk, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
ShlfBrk_threshold <- ShlfBrk.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.ShlfBrk = ShlfBrk.Score) %>%
  select(CellID_2km, Score.ShlfBrk)

ShlfBrk_scored <- grid.2km %>%
  inner_join(ShlfBrk_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.ShlfBrk) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
ShlfBrk_scored_long <- pivot_longer(
  ShlfBrk_scored,
  cols = starts_with(c("0", "1")),
  names_to = "ShlfBrk",
  values_to = "Score.ShlfBrk"
) %>%
  sf::st_transform(4326) %>%
  select(-ShlfBrk) %>%
  st_zm()

st_write_parquet(
  ShlfBrk_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\ShlfBrk_scored_2km_new.parquet"
)
```


# EFHCA - Essential Fish Habitat Conservation Areas

```{r efhca}
# read file in
EFHCA <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_ConsArea_diss"
) %>%
  st_transform(crsOut)

EFHCA.Score = 0.1

# intersect and calculate the area of the overlapping pieces
EFHCA.grid <- sf::st_intersection(EFHCA, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
EFHCA_threshold <- EFHCA.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.EFHCA = EFHCA.Score) %>%
  select(CellID_2km, Score.EFHCA)

EFHCA_scored <- grid.2km %>%
  inner_join(EFHCA_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.EFHCA) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
EFHCA_scored_long <- pivot_longer(
  EFHCA_scored,
  cols = starts_with(c("0", "1")),
  names_to = "EFHCA",
  values_to = "Score.EFHCA"
) %>%
  sf::st_transform(4326) %>%
  select(-EFHCA) %>%
  st_zm()

st_write_parquet(
  EFHCA_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\EFHCA_scored_2km_new.parquet"
)
```


# EFHCA 700 Fathoms - Essential Fish Habitat Conservation Areas

```{r efhca 700}
# read file in
EFHCA_700 <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_ConsArea_700fm"
) %>%
  st_transform(crsOut)

EFHCA.700.Score = 0.1

# intersect and calculate the area of the overlapping pieces
EFHCA.700.grid <- sf::st_intersection(EFHCA_700, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
EFHCA_700_threshold <- EFHCA.700.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.EFHCA.700 = EFHCA.700.Score) %>%
  select(CellID_2km, Score.EFHCA.700)

EFHCA_700_scored <- grid.2km %>%
  inner_join(EFHCA_700_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.EFHCA.700) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
EFHCA_700_scored_long <- pivot_longer(
  EFHCA_700_scored,
  cols = starts_with(c("0", "1")),
  names_to = "EFHCA_700",
  values_to = "Score.EFHCA.700"
) %>%
  sf::st_transform(4326) %>%
  select(-EFHCA_700) %>%
  st_zm()

st_write_parquet(
  EFHCA_700_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\EFHCA_700_scored_2km_new.parquet"
)
```


# HAPCaoi - Habitat Areas of Particular Concern

```{r HAPCaoi}
# read file in
HAPCaoi <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_HAPC_AOI"
) %>%
  st_transform(crsOut)

HAPC.AOI.Score = 0.1

# intersect and calculate the area of the overlapping pieces
HAPC.AOI.grid <- sf::st_intersection(HAPCaoi, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
HAPCaoi_threshold <- HAPC.AOI.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.HAPC.AOI = HAPC.AOI.Score) %>%
  select(CellID_2km, Score.HAPC.AOI)

HAPCaoi_scored <- grid.2km %>%
  inner_join(HAPCaoi_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.HAPC.AOI) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
HAPCaoi_scored_long <- pivot_longer(
  HAPCaoi_scored,
  cols = starts_with(c("0", "1")),
  names_to = "HAPCaoi",
  values_to = "Score.HAPC.AOI"
) %>%
  sf::st_transform(4326) %>%
  select(-HAPCaoi) %>%
  st_zm()

st_write_parquet(
  HAPCaoi_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\HAPCaoi_scored_2km_new.parquet"
)
```


# HAPCreef - Habitat Areas of Particular Concern Rocky Reef

```{r HAPCreef}
# read file in
HAPCreef <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_HAPC_RockyReef_v20250627"
) %>%
  st_transform(crsOut)

HAPC.AOI.Score = 0.1

# intersect and calculate the area of the overlapping pieces
HAPC.AOI.grid <- sf::st_intersection(HAPCaoi, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
HAPCaoi_threshold <- HAPC.AOI.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.HAPC.AOI = HAPC.AOI.Score) %>%
  select(CellID_2km, Score.HAPC.AOI)

HAPCaoi_scored <- grid.2km %>%
  inner_join(HAPCaoi_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.HAPC.AOI) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
HAPCaoi_scored_long <- pivot_longer(
  HAPCaoi_scored,
  cols = starts_with(c("0", "1")),
  names_to = "HAPCaoi",
  values_to = "Score.HAPC.AOI"
) %>%
  sf::st_transform(4326) %>%
  select(-HAPCaoi) %>%
  st_zm()

st_write_parquet(
  HAPCaoi_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\HAPCaoi_scored_2km_new.parquet"
)
```