---
title: "Full Model Report"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      brand: _brand.yml
    toc: true
    toc_float: true
    self_contained: true
params:
  submodels_used: NULL
  submodel_weights_original: NULL
  natural_resources_components: NULL
  industry_operations_components: NULL
  fisheries_components: NULL
  component_layer_details: NULL
  combined_maps_data: NULL
  full_map: NULL
  data_timestamps: NULL
  aoi_data: NULL
  weight_natural_resources: NULL
  weight_fisheries: NULL
  weight_industry_operations: NULL
---

```{css, include = FALSE}
<style>
body {
max-width: 1200px;
margin: 0 auto;
padding: 20px;
font-family: 'Arial', sans-serif;
}

.main-container {
max-width: 1200px;
margin: 0 auto;
}

h1, h2, h3, h4, h5, h6 {
text-align: center;
margin: 20px 0;
}

.leaflet-container {
margin: 20px auto;
display: block;
}

table {
margin: 20px auto;
max-width: 800px;
}

.table {
margin: 20px auto;
}

p {
text-align: justify;
margin: 15px 0;
}

.tocify-wrapper {
position: fixed;
top: 0;
left: 0;
width: 260px;
height: 100%;
overflow: auto;
z-index: 1000;
}

.col-md-9 {
margin-left: 280px;
max-width: calc(100% - 300px);
}

/* Center specific elements */
.centered-content {
text-align: center;
margin: 20px auto;
}

/* Map container centering */
.map-container {
display: flex;
justify-content: center;
margin: 20px 0;
}

/* Ensure leaflet maps are responsive and centered */
.leaflet {
max-width: 100%;
height: 400px;
margin: 0 auto;
}

/* Style for configuration table */
.config-table {
margin: 20px auto;
max-width: 800px;
}

/* Center all chunk outputs */
.chunk-output {
text-align: center;
}

/* Center kable tables */
.kable-table {
margin-left: auto;
margin-right: auto;
}

/* Submodel summary styling */
.submodel-summary {
background-color: #f8f9fa;
border: 1px solid #dee2e6;
border-radius: 8px;
padding: 15px;
margin: 15px 0;
}

.submodel-header {
background-color: #e9ecef;
border-radius: 5px;
padding: 10px;
margin-bottom: 10px;
font-weight: bold;
}

/* Component summary styling */
.component-summary {
background-color: #f1f8ff;
border: 1px solid #c6e2ff;
border-radius: 5px;
padding: 10px;
margin: 10px 0;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

source("global_report_generation.R")
```

## Model Configuration

```{r, results='asis'}

if (!is.null(params$component_layer_details) && length(params$component_layer_details) > 0) {
  
  # Get data timestamps
  data_timestamps_df <- NULL
  if (!is.null(params$data_timestamps)) {
    data_timestamps_df <- params$data_timestamps %>%
     select(dataset_name, data_type, formatted_date) %>%
     rename(Layer_Name = dataset_name, Data_Type = data_type, Last_Updated = formatted_date)
  }
  
  layer_details_list <- list()
  
  # Iterate through all submodels/components to build the table
  for (submodel in names(params$component_layer_details)) {
    submodel_data <- params$component_layer_details[[submodel]]
    
    if (!is.null(submodel_data) && length(submodel_data) > 0) {
      for (component in names(submodel_data)) {
        component_data <- submodel_data[[component]]
        component_method <- component_data$method %||% "Unknown"
        component_layers <- component_data$layers %||% list()
        
        if (length(component_layers) > 0) {
          for (i in seq_along(component_layers)) {
            layer_details_list[[length(layer_details_list) + 1]] <- data.frame(
              Submodel_Key = submodel, # Keep internal key for weight lookup
              Submodel_Display = case_when(
                submodel == "natural_resources" ~ "Natural Resources",
                submodel == "industry_operations" ~ "Industry & Operations",
                submodel == "fisheries" ~ "Fisheries",
                TRUE ~ submodel
              ),
              Layer_Name = component_layers[[i]]$layer_name %||% "Unknown",
              Score_Used = component_layers[[i]]$score_used %||% "Unknown",
              Component = component,
              Method = component_method,
              stringsAsFactors = FALSE
            )
          }
        }
      }
    }
  }
  
  # 2. GENERATE TABLES IF DATA EXISTS
  if (length(layer_details_list) > 0) {
    all_layer_details <- do.call(rbind, layer_details_list)
    
    # Define Sorting Orders
    submodel_order_keys <- c("natural_resources", "fisheries", "industry_operations")
    component_sort_order <- c("Habitat", "Species", "Fisheries", "Trawl Fisheries", "Scientific Surveys", "Submarine Cables")
    
    # Iterate through each submodel to create separate tables
    for (sm_key in submodel_order_keys) {
      
      # Filter data for this specific submodel
      sm_data <- all_layer_details %>% filter(Submodel_Key == sm_key)
      
      if (nrow(sm_data) > 0) {
        # Get Weight
        weight_val <- params$submodel_weights_original[[sm_key]] %||% 0
        display_name <- unique(sm_data$Submodel_Display)[1]
        
        # Sort the data specifically for this table
        sm_data_sorted <- sm_data %>%
          mutate(sort_id = match(Component, component_sort_order)) %>%
          arrange(sort_id, Layer_Name)
        
        final_table_data <- sm_data_sorted %>%
          select(Layer_Name, Score_Used, Component, Method)
          
        if (!is.null(data_timestamps_df)) {
          final_table_data <- final_table_data %>%
            left_join(data_timestamps_df, by = "Layer_Name") %>%
            distinct(Layer_Name, Component, .keep_all = TRUE) %>%
            mutate(
              Data_Type = ifelse(is.na(Data_Type), "Unknown", Data_Type),
              Last_Updated = ifelse(is.na(Last_Updated), "Unknown", Last_Updated)
            )
            
           # Define columns with timestamps
           col_names <- c("Layer Name", "Selected Score", "Component", "Component Calculation Method", "Data Type", "Last Updated")
        } else {
           # Define columns without timestamps
           col_names <- c("Layer Name", "Selected Score", "Component", "Component Calculation Method")
        }
        
        # Print Header with Weight
        cat(paste0("\n### ", display_name, " (Weight: ", round(weight_val, 2), ")\n\n"))
        
        # Print Table
        print(kable(final_table_data,
              col.names = col_names,
              format = "html",
              table.attr = "class='table table-striped config-table'"))
        
        cat("\n<br>\n") # Add spacing between tables
      }
    }
    
  } else {
    cat("No individual layer details available.\n")
  }
} else {
  cat("Individual layer information not available.\n")
}
```

## Full Combined Model Map

The full combined model represents the integration of all selected submodels using their specified weights. The final combination is calculated using the weighted geometric mean of the submodel scores to produce a unified assessment across all selected domains. Scores are distributed along a 0 to 1 scale. Values closer to 0 indicate a less suitable grid cell for offshore wind energy and values closer to 1 indicate a cell that is more suitable for offshore wind energy. 

```{r, results='asis'}
if (!is.null(params$combined_maps_data)) {
  # Check if combined_data has the geometry and expected columns
  if (inherits(params$combined_maps_data, 'sf')) {
    cat("**Full Model Data Points:** ", nrow(params$combined_maps_data), "  \n")
    cat("**Final Combination Method:** Weighted Geometric Mean of Submodel Scores  \n")
    
    # Show value column information
    value_column <- if("Overall_Geo_mean" %in% names(params$combined_maps_data)) {
      "Overall_Geo_mean"
    } else if("Geo_mean" %in% names(params$combined_maps_data)) {
      "Geo_mean"
    } else {
      "Combined_Score"
    }
    
    if (value_column %in% names(params$combined_maps_data)) {
      valid_values <- params$combined_maps_data[[value_column]][!is.na(params$combined_maps_data[[value_column]])]
      if(length(valid_values) > 0) {
        cat("**Value Range:** ", round(min(valid_values), 3), " to ", round(max(valid_values), 3), "  \n")
      }
    }
    cat("\n")
  }
} else {
  cat("*Full model data not available.*\n\n")
}

if (!is.null(params$full_map)) {
  params$full_map
} else {
  cat("*Full model map not available.*\n")
}
```

