---
title: "`r paste(params$component_name)` Component of the `r paste(params$tab_name, 'Submodel Report')`"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      brand: _brand.yml
    toc: true
    toc_float: true
    self_contained: true
params:
  map_configs: NULL
  combined_data_list: NULL    
  selected_methods: NULL    
  tab_name: NULL
  combined_map_title: "Combined Map"
  data_timestamps: NULL
  component_name: NULL
  aoi_data: NULL
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load only essential libraries for report generation
library(leaflet)
library(sf)
library(dplyr)
library(knitr)
library(htmltools)
library(htmlwidgets)

# NO SOURCING OF GLOBAL.R - USE DATA ALREADY LOADED IN APP
# Source only essential functions directly - use the correct environment
if (file.exists("R/create_individual_map.R")) {
  source("R/create_individual_map.R", local = environment())
  cat("✓ create_individual_map.R sourced successfully\n")
} else {
  cat("✗ create_individual_map.R not found\n")
}

if (file.exists("R/create_combined_map.R")) {
  source("R/create_combined_map.R", local = environment()) 
  cat("✓ create_combined_map.R sourced successfully\n")
} else {
  cat("✗ create_combined_map.R not found\n")
}

if (file.exists("R/create_continuous_palette.R")) {
  source("R/create_continuous_palette.R", local = environment())
  cat("✓ create_continuous_palette.R sourced successfully\n")
} else {
  cat("✗ create_continuous_palette.R not found\n")
}

if (file.exists("R/filter_by_score.R")) {
  source("R/filter_by_score.R", local = environment())
  cat("✓ filter_by_score.R sourced successfully\n")
} else {
  cat("✗ filter_by_score.R not found\n")
}

# Get essential data objects directly from global environment
score_colors <- list(
  "0" = "#000000",     # black
  "0.01" =  "#0D5359", #turquoise   
  "0.001" = "#91CB3E", # lime
  "0.1" = "#E41A1C",  # red
  "0.2" = "#03045E",  # blue
  "0.3" = "#08A045",  # green
  "0.4" = "#805D93",  # purple
  "0.5" = "#EC4E20",  # orange
  "0.6" = "#FEEA00",  # yellow
  "0.7" = "#540b0E",  # brown
  "0.8" = "#F49FBC",  # pink
  "0.9" = "#791E94",  # bright purple 
  "1" = "#BE3E82" # berry
)

# Access the data objects that are already loaded in the app environment
# Use get() with proper error handling
safe_get <- function(obj_name, env = .GlobalEnv) {
  if (exists(obj_name, envir = env)) {
    cat("✓ Found", obj_name, "in global environment\n")
    return(get(obj_name, envir = env))
  } else {
    cat("✗ Missing", obj_name, "in global environment\n")
    return(NULL)
  }
}

# Debug: Check what's available in global environment
cat("\n=== DEBUGGING GLOBAL ENVIRONMENT ===\n")
cat("Objects in .GlobalEnv:\n")
global_objects <- ls(envir = .GlobalEnv)
cat(paste(global_objects, collapse = ", "), "\n")

# Get all the datasets using the safe_get function
canyon <- safe_get("canyon")
ASH <- safe_get("ASH")
ASH_ranked_importance <- safe_get("ASH_ranked_importance")
SSH <- safe_get("SSH")
SSH_ranked_importance <- safe_get("SSH_ranked_importance")
GFBT <- safe_get("GFBT")
GFBT_ranked_importance <- safe_get("GFBT_ranked_importance")
GFP <- safe_get("GFP")
GFP_ranked_importance <- safe_get("GFP_ranked_importance")
GFLL <- safe_get("GFLL")
GFLL_ranked_importance <- safe_get("GFLL_ranked_importance")
grid_test <- safe_get("grid_test")

# Test if the functions are available
cat("\n=== FUNCTION AVAILABILITY CHECKS ===\n")
if (exists("create_individual_map", envir = environment())) {
  cat("✓ create_individual_map function is available\n")
} else {
  cat("✗ create_individual_map function is NOT available\n")
}

if (exists("create_combined_map", envir = environment())) {
  cat("✓ create_combined_map function is available\n")
} else {
  cat("✗ create_combined_map function is NOT available\n")
}

# Debug: Check parameters passed to the report
cat("\n=== PARAMETER DEBUGGING ===\n")
cat("map_configs is null:", is.null(params$map_configs), "\n")
if (!is.null(params$map_configs)) {
  cat("Number of map configs:", length(params$map_configs), "\n")
  for(i in 1:min(3, length(params$map_configs))) {
    config <- params$map_configs[[i]]
    cat("Config", i, "- Layer:", config$layer, "Score:", config$score, "\n")
    cat("Config", i, "- Data is null:", is.null(config$data), "\n")
    if (!is.null(config$data)) {
      cat("Config", i, "- Data rows:", nrow(config$data), "\n")
    }
  }
}

cat("combined_data_list is null:", is.null(params$combined_data_list), "\n")
if (!is.null(params$combined_data_list)) {
  cat("Combined data list length:", length(params$combined_data_list), "\n")
  cat("Combined data list names:", names(params$combined_data_list), "\n")
}

cat("aoi_data is null:", is.null(params$aoi_data), "\n")
if (!is.null(params$aoi_data)) {
  cat("AOI data rows:", nrow(params$aoi_data), "\n")
}
```

## Selected Layers Data Information

This report includes data from the following selected layers:

```{r}
if (is.null(params$data_timestamps)) {
  # Call the same function used in your app
  timestamp_info <- get_data_timestamps()
  data_timestamps_table <- timestamp_info$data_timestamps
} else {
  data_timestamps_table <- params$data_timestamps$data_timestamps
}

if (!is.null(data_timestamps_table)) {
  display_table <- data_timestamps_table %>%
    filter(!is.na(formatted_date)) %>%
    select(dataset_name, description, data_type, formatted_date) %>%
    rename(
      "Layer Name" = dataset_name,
      "Description" = description,
      "Data Type" = data_type,
      "Last Updated" = formatted_date
    )
  
  if (nrow(display_table) > 0) {
    kable(display_table, 
          format = "html",
          table.attr = "class='table table-striped'")
  } else {
    cat("No timestamp information available for the selected data layers.")
  }
} else {
  cat("Timestamp information not available.")
}
```

## Mapped Individual Layers 

The following maps are individual data layers that have been visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: layer name, score assigned, and the number of cells that are represented by this data layer.

```{r individual-maps, results='asis'}

if(!is.null(params$map_configs)) {
  
  for(i in seq_along(params$map_configs)) {
    config <- params$map_configs[[i]]
    
    # Text/Markdown Output must use cat() with results='asis'
    cat("\n\n### Map", i, ":", config$layer, "\n\n")
    cat("**Score:** ", config$score, "  \n")
    
    # Check if data is present
    data_points <- tryCatch(nrow(config$data), error = function(e) 0)
    cat("**Data Points:** ", data_points, "  \n\n")
    
    if (data_points == 0) {
      cat("--- **Map skipped: Input data has 0 rows.** ---\n\n")
      # Print a simple placeholder map
      placeholder <- leaflet() %>% 
        addTiles() %>% 
        addControl(paste("Map skipped: No data for", config$layer), position = "topleft")
      print(htmltools::tagList(placeholder))
      next # Skip map generation
    }

    tryCatch({
      map_result <- create_individual_map(config, params$aoi_data)
      
      if (inherits(map_result, "leaflet")) {
        
        # --- CRITICAL FIX: Force map redraw/initialization using onRender ---
        map_result <- map_result %>% 
          htmlwidgets::onRender(
            "function(el, x) { 
               // Force Leaflet to check its container size after the HTML loads
               setTimeout(function() { 
                 var map = this; 
                 map.invalidateSize(); 
               }, 100); 
             }"
          )
        
        # FIX: Ensure the Leaflet widget is printed inside a fixed-size container.
        print(
          htmltools::tagList(
            htmltools::div(
              map_result, 
              style = "height: 500px; width: 100%; border: 1px solid #ddd; margin-bottom: 20px;"
            )
          )
        )
        
      } else {
        # Fallback to a clear error message in the report output
        cat(paste0("--- **ERROR: Map generation failed for ", config$layer, ". Object is not a Leaflet object.** ---\n\n"))
        
        # Simple placeholder map
        placeholder <- leaflet() %>% 
          addTiles() %>% 
          addControl(paste("Map generation failed for", config$layer), position = "topleft")
        print(htmltools::tagList(placeholder))
      }
      
    }, error = function(e) {
      cat(paste0("\n\n--- **FATAL R ERROR creating map for ", config$layer, ":** ", e$message, " ---\n\n"))
      
      # Create error placeholder map
      error_map <- leaflet() %>% 
        addTiles() %>% 
        addControl(paste("Map generation failed (R Error):", e$message), position = "topleft")
      print(htmltools::tagList(error_map))
    })
    
    cat("\n\n")
  }
} else {
  cat("ERROR: No map configurations available - params$map_configs is NULL\n")
}
```

## Combined Maps

The following maps are combined data layers that have been calculated and visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: the number of cells that are represented by this data layer, and the calculation method used (geometric mean, lowest value, or product method).

```{r combined-maps, results='asis'}

if(!is.null(params$combined_data_list) && length(params$combined_data_list) > 0) {
  for(method in names(params$combined_data_list)) {
    combined_data <- params$combined_data_list[[method]]
    
    # Text/Markdown Output must use cat() with results='asis'
    cat("\n\n### Combined Map -", tools::toTitleCase(gsub("_", " ", method)), "\n\n")

    # Check if data is present
    data_points <- tryCatch(nrow(combined_data), error = function(e) 0)
    cat("**Data Points:** ", data_points, "  \n\n")
    
    if (data_points == 0) {
      cat(paste0("--- **Map skipped: Input data (", method, ") has 0 rows.** ---\n\n"))
       # Print a simple placeholder map
      placeholder <- leaflet() %>% 
        addTiles() %>% 
        addControl(paste("Map skipped: No data for", method), position = "topleft")
      print(htmltools::tagList(placeholder))
      next # Skip map generation
    }
    
    # Create the combined map with error handling
    tryCatch({
      map_result <- create_combined_map(combined_data, 
                                        paste(params$combined_map_title, "-", method), 
                                        method, 
                                        aoi_data = params$aoi_data)
      
      if (inherits(map_result, "leaflet")) {

        # --- CRITICAL FIX: Force map redraw/initialization using onRender ---
        map_result <- map_result %>% 
          htmlwidgets::onRender(
            "function(el, x) { 
               // Force Leaflet to check its container size after the HTML loads
               setTimeout(function() { 
                 var map = this; 
                 map.invalidateSize(); 
               }, 100); 
             }"
          )
          
        # FIX: Ensure the Leaflet widget is printed inside a fixed-size container.
        print(
          htmltools::tagList(
            htmltools::div(
              map_result, 
              style = "height: 500px; width: 100%; border: 1px solid #ddd; margin-bottom: 20px;"
            )
          )
        )

      } else {
        # Fallback to a clear error message in the report output
        cat(paste0("--- **ERROR: Combined Map generation failed for ", method, ". Object is not a Leaflet object.** ---\n\n"))
      }
      
    }, error = function(e) {
      cat(paste0("\n\n--- **FATAL R ERROR creating combined map for ", method, ":** ", e$message, " ---\n\n"))
      
      error_map <- leaflet() %>% 
        addTiles() %>% 
        addControl(paste("Combined map generation failed (R Error):", e$message), position = "topleft")
      print(htmltools::tagList(error_map))
    })
    
    cat("\n\n")
  }
} else {
  cat("No combined map data available.\n")
}

```
