---
title: "`r paste(params$component_name)` Component of the `r paste(params$tab_name, 'Submodel Report')`"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      brand: _brand.yml
    toc: true
    toc_float: true
    self_contained: true
params:
  active_resolution: NULL
  map_configs: NULL
  combined_data_list: NULL    
  selected_methods: NULL    
  tab_name: NULL
  combined_map_title: "Combined Map"
  combined_map_objects: NULL
  data_timestamps: NULL
  component_name: NULL
  aoi_data: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

source("global_report_generation.R")

# Function to render individual maps using function directly
render_individual_map_from_config <- function(config) {
  # Get AOI data same way as the server
  aoi_data <- params$aoi_data

  # Call the pure function directly - this generates the exact same map as the app
  return(create_individual_map(config, aoi_data))
}
```

## Selected Layers Data Information

This report includes data from the following selected layers at the `r params$active_resolution` grid size:

```{r data-time}

if (is.null(params$data_timestamps)) {
  timestamp_info <- get_data_timestamps()
  data_timestamps_table <- timestamp_info$data_timestamps
} else {
  data_timestamps_table <- params$data_timestamps$data_timestamps
}


selected_config_list <- list()

if (!is.null(params$map_configs) && length(params$map_configs) > 0) {
  for (i in seq_along(params$map_configs)) {
    config <- params$map_configs[[i]]

    layer_name <- config$layer %||% "Unknown"
    score_val <- config$score %||% "N/A"

    selected_config_list[[length(selected_config_list) + 1]] <- data.frame(
      dataset_name = layer_name,
      selected_score = score_val,
      order_id = i, # Capture the original order from the app
      stringsAsFactors = FALSE
    )
  }
}

if (length(selected_config_list) > 0) {
  selected_config_df <- do.call(rbind, selected_config_list)
} else {
  selected_config_df <- NULL
}


if (!is.null(data_timestamps_table) && !is.null(selected_config_df)) {
  display_table <- selected_config_df %>%
    left_join(data_timestamps_table, by = "dataset_name") %>%

    distinct(dataset_name, selected_score, .keep_all = TRUE) %>%

    arrange(order_id) %>%

    select(
      dataset_name,
      selected_score,
      description,
      data_type,
      formatted_date
    ) %>%
    rename(
      "Layer Name" = dataset_name,
      "Selected Score" = selected_score,
      "Description" = description,
      "Data Type" = data_type
    )

  # Dynamically rename the timestamp column based on the grid resolution
  names(display_table)[names(display_table) == "formatted_date"] <- paste0(
    "Last Updated (",
    params$active_resolution,
    ")"
  )

  if (nrow(display_table) > 0) {
    kable(
      display_table,
      format = "html",
      table.attr = "class='table table-striped'"
    )
  } else {
    cat("No information available for the selected data layers.")
  }
} else {
  cat("Timestamp or configuration information not available.")
}
```

## Mapped Individual Layers 

The following maps are individual data layers that have been visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: layer name, score assigned, and the number of cells that are represented by this data layer.

```{r individual-map1, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 1) {
  config <- params$map_configs[[1]]

  cat("\n\n### Map", 1, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map2, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 2) {
  config <- params$map_configs[[2]]

  cat("\n\n### Map", 2, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map3, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 3) {
  config <- params$map_configs[[3]]

  cat("\n\n### Map", 3, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map4, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 4) {
  config <- params$map_configs[[4]]

  cat("\n\n### Map", 4, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map5, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 5) {
  config <- params$map_configs[[5]]

  cat("\n\n### Map", 5, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map6, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 6) {
  config <- params$map_configs[[6]]

  cat("\n\n### Map", 6, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map7, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 7) {
  config <- params$map_configs[[7]]

  cat("\n\n### Map", 7, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map8, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 8) {
  config <- params$map_configs[[8]]

  cat("\n\n### Map", 8, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

```{r individual-map9, results='asis'}
if (!is.null(params$map_configs) && length(params$map_configs) >= 9) {
  config <- params$map_configs[[9]]

  cat("\n\n### Map", 9, ":", config$layer, "(Score:", config$score, ")\n\n")

  render_individual_map_from_config(config)
}
```

## Combined Maps

The following maps are combined data layers that have been calculated and visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: the number of cells that are represented by this data layer, and the calculation method used (geometric mean, lowest value, or product method).

```{r geometric-mean-map, results='asis'}
if (
  "geometric_mean" %in%
    params$selected_methods &&
    !is.null(params$combined_map_objects[["geometric_mean"]])
) {
  cat("\n\n### Combined Map - Geometric Mean\n\n")

  # Add data summary information
  if (!is.null(params$combined_data_list[["geometric_mean"]])) {
    combined_data <- params$combined_data_list[["geometric_mean"]]
    if (inherits(combined_data, 'sf')) {
      cat("**Data Points:** ", nrow(combined_data), "  \n")

      if ("Geo_mean" %in% names(combined_data)) {
        valid_values <- combined_data[["Geo_mean"]][
          !is.na(combined_data[["Geo_mean"]])
        ]
        if (length(valid_values) > 0) {
          cat(
            "**Value Range:** ",
            round(min(valid_values), 3),
            " to ",
            round(max(valid_values), 3),
            "  \n"
          )
        }
      }
      cat("**Calculation Method:** Geometric Mean  \n")
      cat("\n")
    }
  }

  # Display the map
  params$combined_map_objects[["geometric_mean"]]
}
```

```{r lowest-map, results='asis'}
if (
  "lowest" %in%
    params$selected_methods &&
    !is.null(params$combined_map_objects[["lowest"]])
) {
  cat("\n\n### Combined Map - Lowest Value\n\n")

  # Add data summary information
  if (!is.null(params$combined_data_list[["lowest"]])) {
    combined_data <- params$combined_data_list[["lowest"]]
    if (inherits(combined_data, 'sf')) {
      cat("**Data Points:** ", nrow(combined_data), "  \n")

      if ("Lowest_value" %in% names(combined_data)) {
        valid_values <- combined_data[["Lowest_value"]][
          !is.na(combined_data[["Lowest_value"]])
        ]
        if (length(valid_values) > 0) {
          cat(
            "**Value Range:** ",
            round(min(valid_values), 3),
            " to ",
            round(max(valid_values), 3),
            "  \n"
          )
        }
      }
      cat("**Calculation Method:** Lowest Value  \n")
      cat("\n")
    }
  }

  # Display the map
  params$combined_map_objects[["lowest"]]
}
```

```{r product-map, results='asis'}
if (
  "product" %in%
    params$selected_methods &&
    !is.null(params$combined_map_objects[["product"]])
) {
  cat("\n\n### Combined Map - Product\n\n")

  # Add data summary information
  if (!is.null(params$combined_data_list[["product"]])) {
    combined_data <- params$combined_data_list[["product"]]
    if (inherits(combined_data, 'sf')) {
      cat("**Data Points:** ", nrow(combined_data), "  \n")

      if ("Product_value" %in% names(combined_data)) {
        valid_values <- combined_data[["Product_value"]][
          !is.na(combined_data[["Product_value"]])
        ]
        if (length(valid_values) > 0) {
          cat(
            "**Value Range:** ",
            round(min(valid_values), 3),
            " to ",
            round(max(valid_values), 3),
            "  \n"
          )
        }
      }
      cat("**Calculation Method:** Product  \n")
      cat("\n")
    }
  }

  # Display the map
  params$combined_map_objects[["product"]]
}
```
