---
title: "`r paste(params$submodel_name, 'Combined Submodel Report')`"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      brand: _brand.yml
    toc: true
    toc_float: true
    self_contained: true
params:
  submodel_name: "Combined Submodel"
  selected_components: NULL
  component_methods: NULL
  combined_data: NULL
  combined_map_title: "Combined Submodel Map"
  data_timestamps: NULL
  aoi_data: NULL
  component_layer_details: NULL
  component_data_summary: NULL
  combined_map: NULL,
  aoi_cropped_map: NULL,
  aoi_cropped_normalized_map: NULL
---

```{css, include = FALSE}
<style>
body {
max-width: 1200px;
margin: 0 auto;
padding: 20px;
font-family: 'Arial', sans-serif;
}

.main-container {
max-width: 1200px;
margin: 0 auto;
}

h1, h2, h3, h4, h5, h6 {
text-align: center;
margin: 20px 0;
}

.leaflet-container {
margin: 20px auto;
display: block;
}

table {
margin: 20px auto;
max-width: 800px;
}

.table {
margin: 20px auto;
}

p {
text-align: justify;
margin: 15px 0;
}

.tocify-wrapper {
position: fixed;
top: 0;
left: 0;
width: 260px;
height: 100%;
overflow: auto;
z-index: 1000;
}

.col-md-9 {
margin-left: 280px;
max-width: calc(100% - 300px);
}

/* Center specific elements */
.centered-content {
text-align: center;
margin: 20px auto;
}

/* Map container centering */
.map-container {
display: flex;
justify-content: center;
margin: 20px 0;
}

/* Ensure leaflet maps are responsive and centered */
.leaflet {
max-width: 100%;
height: 400px;
margin: 0 auto;
}

/* Style for configuration table */
.config-table {
margin: 20px auto;
max-width: 800px;
}

/* Center all chunk outputs */
.chunk-output {
text-align: center;
}

/* Center kable tables */
.kable-table {
margin-left: auto;
margin-right: auto;
}

/* Component summary styling */
.component-summary {
background-color: #f8f9fa;
border: 1px solid #dee2e6;
border-radius: 8px;
padding: 15px;
margin: 15px 0;
}

.component-header {
background-color: #e9ecef;
border-radius: 5px;
padding: 10px;
margin-bottom: 10px;
font-weight: bold;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 8)

source("global.R")

# Null coalescing operator
`%||%` <- function(x, y) if(is.null(x)) y else x

```

## Selected Data Layer Information
```{r}
# Add detailed component layer information
if (!is.null(params$component_layer_details) && length(params$component_layer_details) > 0) {
  cat("Individual Layers Used in Combined Submodel\n\n")
  
  # Create a comprehensive table showing all layers used in each component
  layer_details_list <- list()
  
  for (component in names(params$component_layer_details)) {
    component_data <- params$component_layer_details[[component]]
    component_method <- component_data$method %||% "Unknown"
    component_layers <- component_data$layers %||% list()
    
    
    if (length(component_layers) > 0) {
      for (i in seq_along(component_layers)) {
        # Safely extract values with fallbacks
        layer_name <- component_layers[[i]]$layer_name %||% "Unknown"
        score_used <- component_layers[[i]]$score_used %||% "Unknown"
        
        layer_details_list[[length(layer_details_list) + 1]] <- data.frame(
          Component = component,
          Method = component_method,
          Layer_Name = layer_name,
          Score_Used = score_used,
          stringsAsFactors = FALSE
        )
      }
    }
  }
  
  if (length(layer_details_list) > 0) {
    # Combine all layer details into one data frame
    all_layer_details <- do.call(rbind, layer_details_list)
    
    kable(all_layer_details, 
          col.names = c("Component", " Component Calculation Method", "Layer Name", "Selected Score"),
          format = "html",
          table.attr = "class='table table-striped config-table'")
  } else {
    cat("No individual layer details available.\n")
  }
} else {
  cat("Individual layer information not available.\n")
}
```

## Combined Submodel

The combined submodel map represents the integration of all selected components using their specified calculation methods. The final combination is calculated using the geometric mean of the component scores to produce a unified assessment.

```{r}
if (!is.null(params$combined_data)) {
  # Check if combined_data has the geometry and expected columns
  if (inherits(params$combined_data, 'sf')) {
    cat("**Combined Submodel Data Points:** ", nrow(params$combined_data), "  \n")
    cat("**Final Combination Method:** Geometric Mean of Component Scores  \n")
    
    # Show value column information
    value_column <- if("Geo_mean" %in% names(params$combined_data)) "Geo_mean" else "Combined_Score"
    if (value_column %in% names(params$combined_data)) {
      valid_values <- params$combined_data[[value_column]][!is.na(params$combined_data[[value_column]])]
      if(length(valid_values) > 0) {
        cat("**Value Range:** ", round(min(valid_values), 3), " to ", round(max(valid_values), 3), "  \n")
      }
    }
    cat("\n")
  }
} else {
  cat("*Combined submodel data not available.*\n\n")
}

if (!is.null(params$combined_map)) {
  params$combined_map
} else {
  cat("*Combined submodel map not available.*\n")
}
```

## Combined Submodel - Cropped to area of interest (AOI)

The AOI-cropped map shows the same combined submodel data filtered to the selected Area of Interest (AOI), maintaining the original color scale for consistency with the full dataset.

```{r}
if (!is.null(params$aoi_cropped_map)) {
  params$aoi_cropped_map
} else {
  cat("*AOI-cropped map not available. Please select a specific AOI area.*\n")
}
```

## Normalized Combined Submodel - Cropped to area of interest (AOI)

he normalized AOI-cropped map shows the AOI-filtered data rescaled to a 0-1 range using a min max normalization for only the values within the selected area, making it easier to compare relative differences within the AOI.

```{r}
if (!is.null(params$aoi_cropped_normalized_map)) {
  params$aoi_cropped_normalized_map
} else {
  cat("*AOI-cropped normalized map not available. Please select a specific AOI area.*\n")
}
```