---
title: "`r paste(params$submodel_name, 'Combined Submodel Report')`"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      brand: _brand.yml
    toc: true
    toc_float: true
    self_contained: true
params:
  active_resolution: NULL
  submodel_name: "Combined Submodel"
  selected_components: NULL
  component_methods: NULL
  combined_data: NULL
  combined_map_title: "Combined Submodel Map"
  data_timestamps: NULL
  aoi_data: NULL
  component_layer_details: NULL
  component_data_summary: NULL
  combined_map: NULL,
  aoi_cropped_normalized_map: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

source("global_report_generation.R")
```

## Selected Data Layer Information

This report includes data from the following selected layers at the `r params$active_resolution` grid size:

```{r}
# Add detailed component layer information with timestamps
if (
  !is.null(params$component_layer_details) &&
    length(params$component_layer_details) > 0
) {
  # Get data timestamps for joining
  data_timestamps_df <- NULL
  if (!is.null(params$data_timestamps)) {
    # grab the raw dataframe
    temp_df <- params$data_timestamps$data_timestamps

    # force the correct resolution column to be named 'formatted_date'
    if (params$active_resolution == "2km") {
      temp_df$formatted_date <- temp_df$formatted_date_2km
    } else {
      temp_df$formatted_date <- temp_df$formatted_date_5km
    }

    # pipe it into the table
    data_timestamps_df <- temp_df |>
      select(dataset_name, data_type, formatted_date) |>
      rename(
        Layer_Name = dataset_name,
        Data_Type = data_type,
        Last_Updated = formatted_date
      )
  }

  # Create a comprehensive table showing all layers used in each component
  layer_details_list <- list()

  # Component order to maintain
  preferred_order <- c(
    "Habitat",
    "Species",
    "Fisheries",
    "Trawl Fisheries",
    "Scientific Surveys",
    "Submarine Cables"
  )

  available_components <- names(params$component_layer_details)

  sorted_components <- intersect(preferred_order, available_components)

  # Append any other components that might exist but weren't in our preferred list
  remaining <- setdiff(available_components, preferred_order)
  final_component_order <- c(sorted_components, remaining)

  global_row_counter <- 1

  for (component in final_component_order) {
    component_data <- params$component_layer_details[[component]]
    component_method <- component_data$method %||% "Unknown"
    component_layers <- component_data$layers %||% list()

    if (length(component_layers) > 0) {
      for (i in seq_along(component_layers)) {
        # Safely extract values with fallbacks
        layer_name <- component_layers[[i]]$layer_name %||% "Unknown"
        score_used <- component_layers[[i]]$score_used %||% "Unknown"

        layer_details_list[[length(layer_details_list) + 1]] <- data.frame(
          Layer_Name = layer_name,
          Score_Used = score_used,
          Component = component,
          Method = component_method,
          order_id = global_row_counter, # Capture explicit order
          stringsAsFactors = FALSE
        )
        global_row_counter <- global_row_counter + 1
      }
    }
  }

  if (length(layer_details_list) > 0) {
    # Combine all layer details into one data frame
    all_layer_details <- do.call(rbind, layer_details_list)

    # Join with timestamps if available
    if (!is.null(data_timestamps_df)) {
      all_layer_details <- all_layer_details |>
        left_join(data_timestamps_df, by = "Layer_Name") |>
        # Remove duplicates if any
        distinct(Layer_Name, Component, .keep_all = TRUE) |>
        arrange(order_id) |>
        mutate(
          Data_Type = ifelse(is.na(Data_Type), "Unknown", Data_Type),
          Last_Updated = ifelse(is.na(Last_Updated), "Unknown", Last_Updated)
        )

      # Display table with timestamps
      kable(
        all_layer_details |> select(-order_id),
        col.names = c(
          "Layer Name",
          "Selected Score",
          "Component",
          "Component Calculation Method",
          "Data Type",
          paste0("Last Updated (", params$active_resolution, ")")
        ),
        format = "html",
        table.attr = "class='table table-striped config-table'"
      )
    } else {
      # Display table without timestamps (still sorted)
      kable(
        all_layer_details %>% select(-order_id),
        col.names = c(
          "Layer Name",
          "Selected Score",
          "Component",
          "Component Calculation Method"
        ),
        format = "html",
        table.attr = "class='table table-striped config-table'"
      )
    }
  } else {
    cat("No individual layer details available.\\n")
  }
} else {
  cat("Individual layer information not available.\\n")
}
```

## Combined Submodel

The combined submodel map represents the integration of all selected components using their specified calculation methods for the selected area of interest. The final combination is calculated using the geometric mean of the component scores to produce a unified assessment. This dataset are the scores that will be passed to the full model calculation.

```{r, results='asis'}
if (!is.null(params$combined_data)) {
  # Check if combined_data has the geometry and expected columns
  if (inherits(params$combined_data, 'sf')) {
    # 1. Data Points (Standardized Label)
    cat("**Data Points:** ", nrow(params$combined_data), "  \n")

    value_column <- if ("Geo_mean" %in% names(params$combined_data)) {
      "Geo_mean"
    } else {
      "Combined_Score"
    }
    if (value_column %in% names(params$combined_data)) {
      valid_values <- params$combined_data[[value_column]][
        !is.na(params$combined_data[[value_column]])
      ]
      if (length(valid_values) > 0) {
        cat(
          "**Value Range:** ",
          round(min(valid_values), 3),
          " to ",
          round(max(valid_values), 3),
          "  \n"
        )
      }
    }

    cat("**Calculation Method:** Geometric Mean  \n")

    cat("\n")
  }
} else {
  cat("*Combined submodel data not available.*")
}

# Display the map
if (!is.null(params$combined_map)) {
  params$combined_map
} else {
  cat("*Combined submodel map not available.*")
}
```

## Normalized Combined Submodel

This map shows the data rescaled to a 0 to 1 range using a min-max normalization for the cells within the selected area of interest. The goal of this map is to visually highlight differences between cells that are closer in score to make it easier to compare relative differences within the area of interest. The values given are **not** used in any downstream calculations, and this map is purely for visual . 

```{r}
if (!is.null(params$aoi_cropped_normalized_map)) {
  params$aoi_cropped_normalized_map
} else {
  cat(
    "*AOI-cropped normalized map not available. Please select a specific AOI area.*"
  )
}
```