---
title: "`r paste(params$component_name)` Component of the `r paste(params$tab_name, 'Submodel Report')`"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      # brand: _brand.yml #removed to test compatibility in future can copy over in global.R
    toc: true
    toc_float: true
    self_contained: true
params:
  map_configs: NULL
  individual_map_objects: NULL
  combined_data_list: NULL    
  selected_methods: NULL    
  tab_name: NULL
  combined_map_title: "Combined Map"
  data_timestamps: NULL
  component_name: NULL
  aoi_data: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load only essential libraries for report generation
library(leaflet)
library(sf)
library(dplyr)
library(knitr)
library(htmltools)
library(htmlwidgets)

# Source only essential functions directly - use the correct environment
if (file.exists("R/create_individual_map.R")) {
  source("R/create_individual_map.R", local = environment())
  cat("✓ create_individual_map.R sourced successfully\n")
} else {
  cat("✗ create_individual_map.R not found\n")
}

if (file.exists("R/create_combined_map.R")) {
  source("R/create_combined_map.R", local = environment()) 
  cat("✓ create_combined_map.R sourced successfully\n")
} else {
  cat("✗ create_combined_map.R not found\n")
}

if (file.exists("R/create_continuous_palette.R")) {
  source("R/create_continuous_palette.R", local = environment())
  cat("✓ create_continuous_palette.R sourced successfully\n")
} else {
  cat("✗ create_continuous_palette.R not found\n")
}

# Get essential data objects directly from global environment
score_colors <- list(
  "0" = "#000000",     # black
  "0.01" =  "#0D5359", #turquoise   
  "0.001" = "#91CB3E", # lime
  "0.1" = "#E41A1C",  # red
  "0.2" = "#03045E",  # blue
  "0.3" = "#08A045",  # green
  "0.4" = "#805D93",  # purple
  "0.5" = "#EC4E20",  # orange
  "0.6" = "#FEEA00",  # yellow
  "0.7" = "#540b0E",  # brown
  "0.8" = "#F49FBC",  # pink
  "0.9" = "#791E94",  # bright purple 
  "1" = "#BE3E82" # berry
)
```

## Selected Layers Data Information

This report includes data from the following selected layers:

```{r}
if (is.null(params$data_timestamps)) {
  # Call the same function used in app
  timestamp_info <- get_data_timestamps()
  data_timestamps_table <- timestamp_info$data_timestamps
} else {
  data_timestamps_table <- params$data_timestamps$data_timestamps
}

if (!is.null(data_timestamps_table)) {
  display_table <- data_timestamps_table %>%
    filter(!is.na(formatted_date)) %>%
    select(dataset_name, description, data_type, formatted_date) %>%
    rename(
      "Layer Name" = dataset_name,
      "Description" = description,
      "Data Type" = data_type,
      "Last Updated" = formatted_date
    )
  
  if (nrow(display_table) > 0) {
    kable(display_table, 
          format = "html",
          table.attr = "class='table table-striped'")
  } else {
    cat("No timestamp information available for the selected data layers.")
  }
} else {
  cat("Timestamp information not available.")
}
```

## Mapped Individual Layers 

The following maps are individual data layers that have been visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: layer name, score assigned, and the number of cells that are represented by this data layer.

```{r individual-maps, results='asis', fig.align='center'}
if(!is.null(params$individual_map_objects)) {
  
  for(i in seq_along(params$individual_map_objects)) {
    map_obj <- params$individual_map_objects[[i]]
    
    cat("\n\n### Map", i, ":", map_obj$layer, "\n\n")
    cat("**Score:** ", map_obj$score, "  \n")
    cat("**Data Points:** ", nrow(map_obj$config$data), "  \n\n")
    
    # Use only the working temporary file approach
    tryCatch({
      map_result <- create_individual_map(map_obj$config, params$aoi_data)
      
      if (inherits(map_result, "leaflet")) {
        temp_file <- tempfile(fileext = ".html")
        htmlwidgets::saveWidget(map_result, temp_file, selfcontained = TRUE)
        cat('<iframe src="', temp_file, '" width="100%" height="500px" frameborder="0"></iframe>\n')
      }
    }, error = function(e) {
      cat("ERROR creating map:", e$message, "\n")
    })
    
    cat("\n\n")
  }
}
```

## Combined Maps

The following maps are combined data layers that have been calculated and visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: the number of cells that are represented by this data layer, and the calculation method used (geometric mean, lowest value, or product method).

```{r combined-maps, results='asis', fig.align='center'}
if(!is.null(params$combined_data_list) && length(params$combined_data_list) > 0) {
  for(method in names(params$combined_data_list)) {
    combined_data <- params$combined_data_list[[method]]
    
    cat("\n\n### Combined Map -", tools::toTitleCase(gsub("_", " ", method)), "\n\n")
    
    # Create the combined map with error handling
    tryCatch({
      map_result <- create_combined_map(combined_data, 
                                       paste(params$combined_map_title, "-", method), 
                                       method, 
                                       aoi_data = params$aoi_data)
      
      if (inherits(map_result, "leaflet")) {
      
        temp_file <- tempfile(fileext = ".html")
        htmlwidgets::saveWidget(map_result, temp_file, selfcontained = TRUE)
        
      
        cat('<iframe src="', temp_file, '" width="100%" height="500px" frameborder="0"></iframe>\n')
        
      } else {
        cat("ERROR: Map result is not a leaflet object\n")
      }
    }, error = function(e) {
      cat("Error creating combined map:", e$message, "\n")
      error_map <- leaflet() %>% 
        addTiles() %>% 
        addControl("Combined map generation failed", position = "center")
      print(htmltools::tagList(error_map))
    })
    
    cat("\n\n")
  }
} else {
  cat("No combined map data available.\n")
}
```