---
title: "`r paste(params$component_name)` Component of the `r paste(params$tab_name, 'Submodel Report')`"
date: "`r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: 
  html_document:
    theme:
      version: 5
      # brand: _brand.yml #removed to test compatibility in future can copy over in global.R
    toc: true
    toc_float: true
    self_contained: true
params:
  map_configs: NULL
  combined_data_list: NULL    
  selected_methods: NULL    
  tab_name: NULL
  combined_map_title: "Combined Map"
  combined_map_objects: NULL
  data_timestamps: NULL
  component_name: NULL
  aoi_data: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

source("global_report_generation.R")

# Function to render individual maps using function directly
render_individual_map_from_config <- function(config) {
  # Get AOI data same way as the server
  aoi_data <- params$aoi_data
  
  # Call the pure function directly - this generates the exact same map as the app
  return(create_individual_map(config, aoi_data))
}
    
```

## Selected Layers Data Information

This report includes data from the following selected layers:

```{r}
if (is.null(params$data_timestamps)) {
  # Call the same function used in app
  timestamp_info <- get_data_timestamps()
  data_timestamps_table <- timestamp_info$data_timestamps
} else {
  data_timestamps_table <- params$data_timestamps$data_timestamps
}

if (!is.null(data_timestamps_table)) {
  display_table <- data_timestamps_table %>%
    filter(!is.na(formatted_date)) %>%
    select(dataset_name, description, data_type, formatted_date) %>%
    rename(
      "Layer Name" = dataset_name,
      "Description" = description,
      "Data Type" = data_type,
      "Last Updated" = formatted_date
    )
  
  if (nrow(display_table) > 0) {
    kable(display_table, 
          format = "html",
          table.attr = "class='table table-striped'")
  } else {
    cat("No timestamp information available for the selected data layers.")
  }
} else {
  cat("Timestamp information not available.")
}
```

## Mapped Individual Layers 

The following maps are individual data layers that have been visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: layer name, score assigned, and the number of cells that are represented by this data layer.

```{r individual-map1, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 1) {
    config <- params$map_configs[[1]]
    
    cat("\n\n### Map", 1, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map2, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 2) {
    config <- params$map_configs[[2]]
    
    cat("\n\n### Map", 2, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map3, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 3) {
    config <- params$map_configs[[3]]
    
    cat("\n\n### Map", 3, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map4, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 4) {
    config <- params$map_configs[[4]]
    
    cat("\n\n### Map", 4, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map5, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 5) {
    config <- params$map_configs[[5]]
    
    cat("\n\n### Map", 5, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map6, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 6) {
    config <- params$map_configs[[6]]
    
    cat("\n\n### Map", 6, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map7, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 7) {
    config <- params$map_configs[[7]]
    
    cat("\n\n### Map", 7, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map8, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 8) {
    config <- params$map_configs[[8]]
    
    cat("\n\n### Map", 8, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

```{r individual-map9, results='asis'}
if(!is.null(params$map_configs) && length(params$map_configs) >= 9) {
    config <- params$map_configs[[9]]
    
    cat("\n\n### Map", 9, ":", config$layer, "(Score:", config$score, ")\n\n")

    render_individual_map_from_config(config)
  }
```

## Combined Maps

The following maps are combined data layers that have been calculated and visualized for the study area. The maps are zoomed to the area of interest that was previously selected, and shown in a red outline. To view the wider study area select the zoom out button on the map. To more closely analyze a cell click on the cell of interest and a pop-up will generate with the assigned score. Above each of the maps is a block containing relevant information: the number of cells that are represented by this data layer, and the calculation method used (geometric mean, lowest value, or product method).

```{r component-combined-maps, results='asis'}
if (!is.null(params$combined_map_objects) && length(params$combined_map_objects) > 0) {
  
  # Display each combined map based on selected methods
  method_display_names <- list(
    geometric_mean = "Geometric Mean",
    lowest = "Lowest Value", 
    product = "Product"
  )
  
  for (method in names(params$combined_map_objects)) {
    combined_data <- params$combined_data_list[[method]]
    combined_map <- params$combined_map_objects[[method]]
    
    # Display method name and description
    method_name <- method_display_names[[method]] %||% method
    cat("\n\n### Combined Map -", method_name, "\n\n")
    
    # Add data summary information
    if (!is.null(combined_data)) {
      if (inherits(combined_data, 'sf')) {
        cat("**Data Points:** ", nrow(combined_data), "  \n")
        
        # Show value range for the specific method
        value_column <- switch(method,
                               "geometric_mean" = "Geo_mean",
                               "lowest" = "Lowest_value", 
                               "product" = "Product_value")
        
        if (!is.null(value_column) && value_column %in% names(combined_data)) {
          valid_values <- combined_data[[value_column]][!is.na(combined_data[[value_column]])]
          if(length(valid_values) > 0) {
            cat("**Value Range:** ", round(min(valid_values), 3), " to ", round(max(valid_values), 3), "  \n")
          }
        }
        cat("**Calculation Method:** ", method_name, "  \n")
        cat("\n")
      }
    }
    
    # Render the map using the same pattern as individual maps
    if (!is.null(combined_map)) {
      # Create a rendering function just like individual maps do
      render_combined_map_from_param <- function(map_obj) {
        return(map_obj)
      }
      
      render_combined_map_from_param(combined_map)
    } else {
      cat("*Combined map not available for this method.*\n\n")
    }
  }
} else {
  cat("*No combined maps have been generated for this component. Generate combined maps in the application first.*\n\n")
}
```