# Data Processing for SMORES binary datasets for the 2km grid 

Methodology: Set threshold - 50%

Plan: each dataset will have its own code chunk. All processing will live in that code chunk. The top of the document will house functions, data pathways, and other pieces that will be utilised throughout the script. 

```{r, set-up}
#load libraries
library(tidyverse)
library(sf)
library(arcgis)
library(arrow)
library(sfarrow)
library(units)

# 2km base grid
grid.2km <- sf::st_read(
  dsn = "G:\\Shared drives\\NMFS NWC FRAM GIS\\GIS_Base\\Grid.gdb",
  layer = "grid2km_ETsquare"
)

# st_write_parquet(
#   grid.2km,
#   "C:\\GitHub\\SMORES\\data\\2km\\grid_full.parquet"
# )

# get base grid as a dataframe
base_grid_df <- grid.2km %>%
  st_drop_geometry()

# write_parquet(
#   base_grid_df,
#   "C:\\GitHub\\SMORES\\data\\5km\\base_grid_df.parquet"
# )

# set standard coordinate reference system if using 2 km square grid
crsOut <- st_crs(grid.2km)

# calculate the area of the grid cells
grid_2km <- grid.2km %>%
  mutate(cell_area = st_area(.))

# file path where collections of layers lives
melissa_file_path <- "G:\\Shared drives\\NMFS NWC OEI GIS\\ArcGIS\\Projects\\OWEC\\p30\\nccos_share_CA2.gdb"

```

# Habitat Layers

## Canyons

```{r canyons}
# read file in
Canyons <- sf::st_read(
  dsn = melissa_file_path,
  layer = "SubmarineCanyons_WestCoast_diss"
) %>%
  st_transform(crsOut)

Canyons.Score = 0.1

# intersect and calculate the area of the overlapping pieces
Canyons.grid <- sf::st_intersection(Canyons, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
canyon_threshold <- Canyons.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.Canyons = Canyons.Score) %>%
  select(CellID_2km, Score.Canyons)

canyon_scored <- grid.2km %>%
  inner_join(canyon_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.Canyons) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
canyons_scored_long <- pivot_longer(
  canyon_scored,
  cols = starts_with(c("0", "1")),
  names_to = "Canyon",
  values_to = "Score.Canyon"
) %>%
  sf::st_transform(4326) %>%
  select(-Canyon) %>%
  st_zm()

st_write_parquet(
  canyons_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\canyon_scored_full.parquet"
)
```

## Deep Sea Coral

```{r deep-sea-coral}
# read file in
DSC.RobustHigh <- sf::st_read(
  dsn = melissa_file_path,
  layer = "DSC_HabitatSuitability_RobustHigh"
) %>%
  st_transform(crsOut)

DSC.RH.Score = 0.1

# intersect and calculate the area of the overlapping pieces
DSC.RobustHigh.grid <- sf::st_intersection(DSC.RobustHigh, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
DSC.RobustHigh_threshold <- DSC.RobustHigh.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.DSC.RH = DSC.RH.Score) %>%
  select(CellID_2km, Score.DSC.RH)

DSC_RH_scored <- grid.2km %>%
  inner_join(DSC.RobustHigh_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.DSC.RH) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
DSC_RH_scored_long <- pivot_longer(
  DSC_RH_scored,
  cols = starts_with(c("0", "1")),
  names_to = "DSC_RH",
  values_to = "Score.DSC_RH"
) %>%
  sf::st_transform(4326) %>%
  select(-DSC_RH) %>%
  st_zm()

st_write_parquet(
  DSC_RH_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\DSC_RH_scored_full.parquet"
)
```


## Seeps

```{r seeps}
# read file in
Seeps <- sf::st_read(
  dsn = melissa_file_path,
  layer = "CascadiaSeeps_wBuffer"
) %>%
  st_transform(crsOut)

Seeps.Score = 0.1

# intersect and calculate the area of the overlapping pieces
Seeps.grid <- sf::st_intersection(Seeps, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
Seeps_threshold <- Seeps.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.Seeps = Seeps.Score) %>%
  select(CellID_2km, Score.Seeps)

Seeps_scored <- grid.2km %>%
  inner_join(Seeps_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.Seeps) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
Seeps_scored_long <- pivot_longer(
  Seeps_scored,
  cols = starts_with(c("0", "1")),
  names_to = "Seeps",
  values_to = "Score.Seeps"
) %>%
  sf::st_transform(4326) %>%
  select(-Seeps) %>%
  st_zm()

st_write_parquet(
  Seeps_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Seeps_scored_full.parquet"
)
```


## Shelf Break

```{r shelf-break}
# read file in
ShlfBrk <- sf::st_read(
  dsn = melissa_file_path,
  layer = "ContinentalShelfBreak_wBuffer"
) %>%
  st_transform(crsOut)

ShlfBrk.Score = 0.1

# intersect and calculate the area of the overlapping pieces
ShlfBrk.grid <- sf::st_intersection(ShlfBrk, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
ShlfBrk_threshold <- ShlfBrk.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.ShlfBrk = ShlfBrk.Score) %>%
  select(CellID_2km, Score.ShlfBrk)

ShlfBrk_scored <- grid.2km %>%
  inner_join(ShlfBrk_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.ShlfBrk) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
ShlfBrk_scored_long <- pivot_longer(
  ShlfBrk_scored,
  cols = starts_with(c("0", "1")),
  names_to = "ShlfBrk",
  values_to = "Score.ShlfBrk"
) %>%
  sf::st_transform(4326) %>%
  select(-ShlfBrk) %>%
  st_zm()

st_write_parquet(
  ShlfBrk_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\ShlfBrk_scored_full.parquet"
)
```


## EFHCA - Essential Fish Habitat Conservation Areas

```{r efhca}
# read file in
EFHCA <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_ConsArea_diss"
) %>%
  st_transform(crsOut)

EFHCA.Score = 0.1

# intersect and calculate the area of the overlapping pieces
EFHCA.grid <- sf::st_intersection(EFHCA, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
EFHCA_threshold <- EFHCA.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.EFHCA = EFHCA.Score) %>%
  select(CellID_2km, Score.EFHCA)

EFHCA_scored <- grid.2km %>%
  inner_join(EFHCA_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.EFHCA) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
EFHCA_scored_long <- pivot_longer(
  EFHCA_scored,
  cols = starts_with(c("0", "1")),
  names_to = "EFHCA",
  values_to = "Score.EFHCA"
) %>%
  sf::st_transform(4326) %>%
  select(-EFHCA) %>%
  st_zm()

st_write_parquet(
  EFHCA_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\EFHCA_scored_full.parquet"
)
```


## EFHCA 700 Fathoms - Essential Fish Habitat Conservation Areas

```{r efhca-700}
# read file in
EFHCA_700 <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_ConsArea_700fm"
) %>%
  st_transform(crsOut)

EFHCA.700.Score = 0.1

# intersect and calculate the area of the overlapping pieces
EFHCA.700.grid <- sf::st_intersection(EFHCA_700, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
EFHCA_700_threshold <- EFHCA.700.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.EFHCA.700 = EFHCA.700.Score) %>%
  select(CellID_2km, Score.EFHCA.700)

EFHCA_700_scored <- grid.2km %>%
  inner_join(EFHCA_700_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.EFHCA.700) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
EFHCA_700_scored_long <- pivot_longer(
  EFHCA_700_scored,
  cols = starts_with(c("0", "1")),
  names_to = "EFHCA_700",
  values_to = "Score.EFHCA.700"
) %>%
  sf::st_transform(4326) %>%
  select(-EFHCA_700) %>%
  st_zm()

st_write_parquet(
  EFHCA_700_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\EFHCA_700_scored_full.parquet"
)
```


## HAPCaoi - Habitat Areas of Particular Concern

```{r HAPCaoi}
# read file in
HAPCaoi <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_HAPC_AOI"
) %>%
  st_transform(crsOut)

HAPC.AOI.Score = 0.1

# intersect and calculate the area of the overlapping pieces
HAPC.AOI.grid <- sf::st_intersection(HAPCaoi, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
HAPCaoi_threshold <- HAPC.AOI.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.HAPC.AOI = HAPC.AOI.Score) %>%
  select(CellID_2km, Score.HAPC.AOI)

HAPCaoi_scored <- grid.2km %>%
  inner_join(HAPCaoi_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.HAPC.AOI) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
HAPCaoi_scored_long <- pivot_longer(
  HAPCaoi_scored,
  cols = starts_with(c("0", "1")),
  names_to = "HAPCaoi",
  values_to = "Score.HAPC.AOI"
) %>%
  sf::st_transform(4326) %>%
  select(-HAPCaoi) %>%
  st_zm()

st_write_parquet(
  HAPCaoi_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\HAPCaoi_scored_full.parquet"
)
```


## HAPCreef - Habitat Areas of Particular Concern Rocky Reef

```{r HAPCreef}

# read file in
HAPCreef <- sf::st_read(
  dsn = melissa_file_path,
  layer = "EFH_Groundfish_HAPC_RockyReef_v20250627_diss"
) %>%
  st_transform(crsOut)

HAPC.RR.Score = 0.1

# intersect and calculate the area of the overlapping pieces
HAPC.RR.grid <- sf::st_intersection(HAPCreef, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
HAPCreef_threshold <- HAPC.RR.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.HAPC.RR = HAPC.RR.Score) %>%
  select(CellID_2km, Score.HAPC.RR)

HAPCreef_scored <- grid.2km %>%
  inner_join(HAPCreef_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.HAPC.RR) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
HAPCreef_scored_long <- pivot_longer(
  HAPCreef_scored,
  cols = starts_with(c("0", "1")),
  names_to = "HAPCreef",
  values_to = "Score.HAPC.RR"
) %>%
  sf::st_transform(4326) %>%
  select(-HAPCreef) %>%
  st_zm()

st_write_parquet(
  HAPCreef_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\HAPCreef_scored_full.parquet"
)
```

# Species Layers


```{r species-set-up}

ESA_Critical_Habitat <- sf::st_read(
  dsn = "G:\\Shared drives\\NMFS NWC OEI GIS\\ArcGIS\\Projects\\OWEC\\p30\\protected_species_density_layers\\nmfs_combined\\Critical_Habitat_Areas_by_the_National_Marine_Fisheries_Service\\Critical_Habitat_Areas_by_the_National_Marine_Fisheries_Service.shp"
)

bia_bw_nms <- "https://services2.arcgis.com/C8EMgrsFcRFL6LrL/ArcGIS/rest/services/BIA_II_WebMerc/FeatureServer/1"
# contains parent and core - parent = entire range, core = intensified use areas such as breeding ground or feeding site
bia_bw_nms_layer <- arc_read(
  bia_bw_nms,
  where = "Common_Name = 'Blue whale'"
) %>%
  st_transform(crsOut)
```

## Killer Whale

```{r killer-whale}
# read file in
killer_whale <- ESA_Critical_Habitat %>%
  filter(commonName == "Whale, killer") %>%
  select(-c(statusType, scientific, listingSta, federalReg)) %>%
  st_transform(crsOut)

killer_whale.Score = 0.1

# intersect and calculate the area of the overlapping pieces
killer_whale.grid <- sf::st_intersection(killer_whale, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
killer_whale_threshold <- killer_whale.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.killer_whale = killer_whale.Score) %>%
  select(CellID_2km, Score.killer_whale)

killer_whale_scored <- grid.2km %>%
  inner_join(killer_whale_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.killer_whale) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
killer_whale_scored_long <- pivot_longer(
  killer_whale_scored,
  cols = starts_with(c("0", "1")),
  names_to = "killer_whale",
  values_to = "Score.killer_whale"
) %>%
  sf::st_transform(4326) %>%
  select(-killer_whale) %>%
  st_zm()

st_write_parquet(
  killer_whale_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Killer_whale_scored_full.parquet"
)
```

## Leatherback Sea Turtle

```{r leatherback-turtle}
# read file in
leatherback_turtle <- ESA_Critical_Habitat %>%
  filter(commonName == "Sea turtle, leatherback") %>%
  select(-c(statusType, scientific, listingSta, federalReg)) %>%
  st_transform(crsOut)

leatherback.Score = 0.1

# intersect and calculate the area of the overlapping pieces
leatherback.grid <- sf::st_intersection(leatherback_turtle, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
leatherback_threshold <- leatherback.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.leatherback = leatherback.Score) %>%
  select(CellID_2km, Score.leatherback)

leatherback_scored <- grid.2km %>%
  inner_join(leatherback_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.leatherback) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
leatherback_scored_long <- pivot_longer(
  leatherback_scored,
  cols = starts_with(c("0", "1")),
  names_to = "leatherback",
  values_to = "Score.leatherback"
) %>%
  sf::st_transform(4326) %>%
  select(-leatherback) %>%
  st_zm()

st_write_parquet(
  leatherback_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Leatherback_turtle_scored_full.parquet"
)
```
## Blue Whale

```{r blue-whale}
# read file in
blue_whale <- bia_bw_nms_layer %>%
  select(c(Common_Name, Stock, Area_Name, geometry)) %>%
  st_transform(crsOut)

blue.Score = 0.1

# intersect and calculate the area of the overlapping pieces
blue.grid <- sf::st_intersection(blue_whale, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
blue_threshold <- blue.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.blue = blue.Score) %>%
  select(CellID_2km, Score.blue)

blue_scored <- grid.2km %>%
  inner_join(blue_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.blue) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
blue_scored_long <- pivot_longer(
  blue_scored,
  cols = starts_with(c("0", "1")),
  names_to = "blue",
  values_to = "Score.blue"
) %>%
  sf::st_transform(4326) %>%
  select(-blue) %>%
  st_zm()

st_write_parquet(
  blue_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Blue_whale_scored_full.parquet"
)
```


## Humpback Whale

```{r}
humpback_whale <- ESA_Critical_Habitat %>%
  filter(commonName == "Whale, humpback") %>%
  select(-c(statusType, scientific, listingSta, federalReg)) %>%
  st_transform(crsOut)

#Error:
# Error in scan(text = lst[[length(lst)]], quiet = TRUE) :
# scan() expected 'a real', got '936246.12987604213.'
# Error in (function (msg)  :
# TopologyException: side location conflict at 92746.754168340936 936246.12987604213. This can occur if the input geometry is invalid.

# returns false for row 5 and 17 of data
# st_is_valid(humpback_whale)

# returns that the two false rows have a self-intersection [92746.7541682779 936246.129876046]
# st_is_valid(humpback_whale, reason = TRUE)

# suggestion is to apply a buffer of 0 to force the creation of a new geometry -> stack overflow article https://stackoverflow.com/questions/66584191/sfst-intersection-virtually-random-error-action
humpback_whale_buffer <- humpback_whale %>%
  st_buffer(0)

# buffer has made all geometries valid
st_is_valid(humpback_whale_buffer, reason = TRUE)

humpback.Score = 0.1

# intersect and calculate the area of the overlapping pieces
humpback.grid <- sf::st_intersection(humpback_whale_buffer, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
humpback_threshold <- humpback.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.humpback = humpback.Score) %>%
  select(CellID_2km, Score.humpback)

humpback_scored <- grid.2km %>%
  inner_join(humpback_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.humpback) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
humpback_scored_long <- pivot_longer(
  humpback_scored,
  cols = starts_with(c("0", "1")),
  names_to = "humpback",
  values_to = "Score.humpback"
) %>%
  sf::st_transform(4326) %>%
  select(-humpback) %>%
  st_zm()

st_write_parquet(
  humpback_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Humpback_whale_scored_full.parquet"
)
```

# Survey Layers

## Fixed Survey Locations

```{r fixed-surveys}
# read file in
Surveys_fixed <- sf::st_read(
  dsn = melissa_file_path,
  layer = "Surveys_Operation_Area_fixed"
) %>%
  st_transform(crsOut)

surveys.fixed.Score = 0.1

# intersect and calculate the area of the overlapping pieces
surveys_fixed.grid <- sf::st_intersection(Surveys_fixed, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
surveys_fixed_threshold <- surveys_fixed.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.surveys.fixed = surveys.fixed.Score) %>%
  select(CellID_2km, Score.surveys.fixed)

surveys_fixed_scored <- grid.2km %>%
  inner_join(surveys_fixed_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.surveys.fixed) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
surveys_fixed_scored_long <- pivot_longer(
  surveys_fixed_scored,
  cols = starts_with(c("0", "1")),
  names_to = "surveys_fixed",
  values_to = "Score.surveys.fixed"
) %>%
  sf::st_transform(4326) %>%
  select(-surveys_fixed) %>%
  st_zm()

st_write_parquet(
  surveys_fixed_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Surveys_fixed_scored_full.parquet"
)
```

## Periodic Survey Locations

```{r}
# read file in
Surveys_periodic <- sf::st_read(
  dsn = melissa_file_path,
  layer = "Surveys_Operation_Area_periodic"
) %>%
  st_transform(crsOut)

surveys.periodic.Score = 0.1

# intersect and calculate the area of the overlapping pieces
surveys_periodic.grid <- sf::st_intersection(Surveys_periodic, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
surveys_periodic_threshold <- surveys_periodic.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.surveys.periodic = surveys.periodic.Score) %>%
  select(CellID_2km, Score.surveys.periodic)

surveys_periodic_scored <- grid.2km %>%
  inner_join(surveys_periodic_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.surveys.periodic) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
surveys_periodic_scored_long <- pivot_longer(
  surveys_periodic_scored,
  cols = starts_with(c("0", "1")),
  names_to = "surveys_periodic",
  values_to = "Score.surveys.periodic"
) %>%
  sf::st_transform(4326) %>%
  select(-surveys_periodic) %>%
  st_zm()

st_write_parquet(
  surveys_periodic_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Surveys_periodic_scored_full.parquet"
)
```

# Cables

## submarine cables

```{r}
furl_nms <- "https://coast.noaa.gov/arcgis/rest/services/Hosted/SubmarineCables/FeatureServer/0"
nms_layer <- arc_read(
  furl_nms,
  where = "status = 'In Service'
                or status = 'Under Construction'"
) %>%
  st_transform(crsOut)

st_is_valid(nms_layer, reason = TRUE)

nms_layer_valid <- st_make_valid(nms_layer)

# cables does not have a great enough width to pass the threshold. Curt requested to know what the width of each # polygon is so we can determine how we want to proceed
cables_width <- nms_layer %>%
  mutate(
    #  area in Transverse Mercator projection (meters^2)
    calc_area = st_area(geometry),
    
    # perimeter by casting the polygon boundary to a line (meters)
    calc_perimeter = st_length(st_cast(geometry, "MULTILINESTRING")),
    
    #  Average Width = (2 * Area) / Perimeter
    avg_width = (2 * calc_area) / calc_perimeter,
    
    avg_width_num = drop_units(avg_width)
  )

# shows that most polygons are between 60-70 m or 0.06-0.07 km
head(cables_width %>% select(avg_width_num))

# Curt has indicated he wants us to apply a 500m buffer and maintain the current 50% threshold
cable_500m_buffered <- st_buffer(cables_width, dist = 500)

# check 
cables_width_buffered <- cable_500m_buffered %>%
  mutate(
    #  area in Transverse Mercator projection (meters^2)
    calc_area = st_area(geometry),
    
    # perimeter by casting the polygon boundary to a line (meters)
    calc_perimeter = st_length(st_cast(geometry, "MULTILINESTRING")),
    
    #  Average Width = (2 * Area) / Perimeter
    avg_width = (2 * calc_area) / calc_perimeter,
    
    avg_width_num = drop_units(avg_width)
  )

head(cables_width_buffered %>% select(avg_width_num))

cable.Score = 0.1

# intersect and calculate the area of the overlapping pieces
cable.grid <- sf::st_intersection(cables_width_buffered, grid_2km) %>%
  mutate(intersect_area = st_area(.))

# sum the overlapping area per cell and calculate the threshold
cable_threshold <- cable.grid %>%
  st_drop_geometry() %>%
  group_by(CellID_2km, cell_area) %>%
  summarize(total_intersect_area = sum(intersect_area), .groups = "drop") %>%
  mutate(
    coverage_pct = as.numeric(total_intersect_area) / as.numeric(cell_area)
  ) %>%
  # keep cells with greater than or equal to 50% coverage
  filter(coverage_pct >= 0.5) %>%
  mutate(Score.cable = cable.Score) %>%
  select(CellID_2km, Score.cable)

cable_scored <- grid.2km %>%
  inner_join(cable_threshold, by = "CellID_2km") %>%
  rename("0.1" = Score.cable) %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001,
    "0.2" = 0.2,
    "0.3" = 0.3,
    "0.4" = 0.4,
    "0.5" = 0.5,
    "0.6" = 0.6,
    "0.7" = 0.7,
    "0.8" = 0.8,
    "0.9" = 0.9,
    "1" = 1
  )

# transform data structure so that score.canyons is filterable by app
cable_scored_long <- pivot_longer(
  cable_scored,
  cols = starts_with(c("0", "1")),
  names_to = "cable",
  values_to = "Score.cable"
) %>%
  sf::st_transform(4326) %>%
  select(-cable) %>%
  st_zm()

st_write_parquet(
  cable_scored_long,
  "C:\\GitHub\\SMORES\\data\\2km\\Submarine_cable_scored_full.parquet"
)
```