# Data Processing for SMORES binary datasets for the 5km grid 

Methodology: Area Weighted Average

Plan: each dataset will have its own code chunk. All processing will live in that code chunk. The top of the document will house functions, data pathways, and other pieces that will be utilised throughout the script. 

```{r setup}
#load libraries
library(tidyverse)
library(sf)
library(arcgis)
library(arrow)
library(sfarrow)
library(ggplot2)

# 2km base grid
grid.5km <- sf::st_read(
  dsn = "G:\\Shared drives\\NMFS NWC FRAM GIS\\GIS_Base\\Grid.gdb",
  layer = "grid5km_ETsquare"
)

# st_write_parquet(
#   grid.2km,
#   "C:\\GitHub\\SMORES\\data\\2km\\grid_full.parquet"
# )

# get base grid as a dataframe
base_grid_df <- grid.5km %>%
  st_drop_geometry()

# write_parquet(
#   base_grid_df,
#   "C:\\GitHub\\SMORES\\data\\5km\\base_grid_df.parquet"
# )

# set standard coordinate reference system if using 2 km square grid
crsOut <- st_crs(grid.5km)

# calculate the area of the grid cells
grid_5km <- grid.5km %>%
  mutate(cell_area = st_area(.))

Fisheries <- sf::st_read(
  "G:\\My Drive\\SMORES\\NCCOSfisherieslayers20250815081621\\Non-Confidential_NMFS_fisheries_submodel_data_2km_grid.shp"
) %>%
  st_transform(crsOut)

Trawl_Fisheries <- sf::st_read(
  "G:\\My Drive\\SMORES\\Scenario_4_trawl_polygon\\Scenario_4_trawl_polygon.shp"
)
```

# Fisheries Layers

## ASH

```{r ASH}
ASH <- Fisheries %>%
  select(ASH_RI)

# calculating the ranked importance scoring
ASH.grid_RI <- sf::st_intersection(ASH, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = ASH_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.ASH_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.ASH_Ranked_Importance)) %>%
  # remove 0 values for continuity with NCCOS report
  filter(Score.ASH_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
ASH_RI_5km <- grid.5km %>%
  inner_join(ASH.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  ASH_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\ASH_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
ASH_presence <- ASH.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

ASH_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(ASH_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "ASH",
    values_to = "Score.ASH"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-ASH)

st_write_parquet(
  ASH_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\ASH_scored_full.parquet"
)
```

## SSH

```{r SSH}
SSH <- Fisheries %>%
  select(SSH_RI)

# calculating the ranked importance scoring
SSH.grid_RI <- sf::st_intersection(SSH, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = SSH_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.SSH_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.SSH_Ranked_Importance)) %>%
  filter(Score.SSH_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
SSH_RI_5km <- grid.5km %>%
  inner_join(SSH.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  SSH_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\SSH_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
SSH_presence <- SSH.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

SSH_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(SSH_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "SSH",
    values_to = "Score.SSH"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-SSH)

st_write_parquet(
  SSH_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\SSH_scored_full.parquet"
)
```

## Groundfish Bottom Trawl - GFBT

```{r GFBT}
GFBT <- Fisheries %>%
  select(GFBT_RI)

# calculating the ranked importance scoring
GFBT.grid_RI <- sf::st_intersection(GFBT, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = GFBT_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.GFBT_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.GFBT_Ranked_Importance)) %>%
  filter(Score.GFBT_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
GFBT_RI_5km <- grid.5km %>%
  inner_join(GFBT.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  GFBT_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\GFBT_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
GFBT_presence <- GFBT.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

GFBT_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(GFBT_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "GFBT",
    values_to = "Score.GFBT"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-GFBT)

st_write_parquet(
  GFBT_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\GFBT_scored_full.parquet"
)
```

## Groundfish Pot Gear - GFP
```{r GFP}
GFP <- Fisheries %>%
  select(GFP_RI)

# calculating the ranked importance scoring
GFP.grid_RI <- sf::st_intersection(GFP, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = GFP_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.GFP_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.GFP_Ranked_Importance)) %>%
  filter(Score.GFP_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
GFP_RI_5km <- grid.5km %>%
  inner_join(GFP.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  GFP_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\GFP_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
GFP_presence <- GFP.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

GFP_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(GFP_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "GFP",
    values_to = "Score.GFP"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-GFP)

st_write_parquet(
  GFP_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\GFP_scored_full.parquet"
)
```

## Groundfish Long Line - GFLL

```{r GFLL}
GFLL <- Fisheries %>%
  select(GFLL_RI)

# calculating the ranked importance scoring
GFLL.grid_RI <- sf::st_intersection(GFLL, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = GFLL_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.GFLL_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.GFLL_Ranked_Importance)) %>%
  filter(Score.GFLL_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
GFLL_RI_5km <- grid.5km %>%
  inner_join(GFLL.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  GFLL_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\GFLL_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
GFLL_presence <- GFLL.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

GFLL_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(GFLL_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "GFLL",
    values_to = "Score.GFLL"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-GFLL)

st_write_parquet(
  GFLL_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\GFLL_scored_full.parquet"
)
```

## Pink Shrimp - PS

```{r PS}
PS <- Fisheries %>%
  select(PS_RI)

# calculating the ranked importance scoring
PS.grid_RI <- sf::st_intersection(PS, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = PS_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.PS_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.PS_Ranked_Importance)) %>%
  filter(Score.PS_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
PS_RI_5km <- grid.5km %>%
  inner_join(PS.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  PS_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\PS_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
PS_presence <- PS.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

PS_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(PS_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "PS",
    values_to = "Score.PS"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-PS)

st_write_parquet(
  PS_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\PS_scored_full.parquet"
)
```

## Dungeness Crab - CRAB

```{r CRAB}
CRAB <- Fisheries %>%
  select(CRAB_RI)

# calculating the ranked importance scoring
CRAB.grid_RI <- sf::st_intersection(CRAB, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = CRAB_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.CRAB_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.CRAB_Ranked_Importance)) %>%
  filter(Score.CRAB_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
CRAB_RI_5km <- grid.5km %>%
  inner_join(CRAB.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  CRAB_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\CRAB_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
CRAB_presence <- CRAB.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

CRAB_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(CRAB_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "CRAB",
    values_to = "Score.CRAB"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-CRAB)

st_write_parquet(
  CRAB_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\CRAB_scored_full.parquet"
)
```

## Commercial troll/hook and line albacore - ALCO

```{r ALCO}
ALCO <- Fisheries %>%
  select(ALCO_RI)

# calculating the ranked importance scoring
ALCO.grid_RI <- sf::st_intersection(ALCO, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = ALCO_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.ALCO_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.ALCO_Ranked_Importance)) %>%
  filter(Score.ALCO_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
ALCO_RI_5km <- grid.5km %>%
  inner_join(ALCO.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  ALCO_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\ALCO_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
ALCO_presence <- ALCO.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

ALCO_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(ALCO_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "ALCO",
    values_to = "Score.ALCO"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-ALCO)

st_write_parquet(
  ALCO_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\ALCO_scored_full.parquet"
)
```

## Charter vessel albacore troll/hook and line - ALCH

```{r ALCH}
ALCH <- Fisheries %>%
  select(ALCH_RI)

# calculating the ranked importance scoring
ALCH.grid_RI <- sf::st_intersection(ALCH, grid_5km) %>%
  mutate(
    # calculate area of each resulting puzzle piece
    intersect_area = sf::st_area(geometry),
    # percentage of the 5km cell this piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area),
    # Multiply the original 2km score by its new spatial weight
    weighted_score = ALCH_RI * area_weight
  ) %>%
  st_drop_geometry() %>%
  group_by(CellID_5km) %>%
  # sum all the weighted pieces together to get the final average
  summarize(
    Score.ALCH_Ranked_Importance = sum(weighted_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(Score.ALCH_Ranked_Importance)) %>%
  filter(Score.ALCH_Ranked_Importance != 0.000)

# join new scores back to 5km spatial grid and transform
ALCH_RI_5km <- grid.5km %>%
  inner_join(ALCH.grid_RI, by = "CellID_5km") %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  ALCH_RI_5km,
  "C:\\GitHub\\SMORES\\data\\5km\\ALCH_Ranked_Importance_scored_full.parquet"
)

# calculating the user selected scores
ALCH_presence <- ALCH.grid_RI %>%
  st_drop_geometry() %>%
  distinct(CellID_5km)

ALCH_scored_static <- grid.5km %>%
  # inner_join automatically drops any 5km cells that don't have ASH present
  inner_join(ALCH_presence, by = "CellID_5km") %>%
  mutate(
    "0" = 0,
    "0.01" = 0.01,
    "0.001" = 0.001
  ) %>%
  pivot_longer(
    cols = c("0", "0.01", "0.001"),
    names_to = "ALCH",
    values_to = "Score.ALCH"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-ALCH)

st_write_parquet(
  ALCH_scored_static,
  "C:\\GitHub\\SMORES\\data\\5km\\ALCH_scored_full.parquet"
)
```

# Deep Sea Coral Robust High 


```{r}
# get base grid as a dataframe
base_grid_df <- grid.2km %>%
  st_drop_geometry()

# write_parquet(
#   base_grid_df,
#   "C:\\GitHub\\SMORES\\data\\5km\\base_grid_df.parquet"
# )

# set standard coordinate reference system if using 2 km square grid
crsOut <- st_crs(grid.2km)
# 2km base grid
grid.2km <- sf::st_read(
  dsn = "G:\\Shared drives\\NMFS NWC FRAM GIS\\GIS_Base\\Grid.gdb",
  layer = "grid2km_ETsquare"
)

# calculate the area of the grid cells
grid_2km <- grid.2km %>%
  mutate(cell_area = st_area(.))


# set standard coordinate reference system if using 2 km square grid
crsOut <- st_crs(grid.2km)

#convert each layer to sf object
melissa_file_path <- "G:\\Shared drives\\NMFS NWC OEI GIS\\ArcGIS\\Projects\\OWEC\\p30\\nccos_share_CA2.gdb"

DSC.RobustHigh <- sf::st_read(
  dsn = melissa_file_path,
  layer = "DSC_HabitatSuitability_RobustHigh"
) %>%
  st_transform(crsOut)

DSC.2km_intersect <- sf::st_intersection(DSC.RobustHigh, grid_2km) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>%
  # keep only the largest overlapping piece, prevent ties from duplicating rows
  slice_max(area.part, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(CellID_2km, gridcode)

# rop geometry to run the Z-membership math efficiently
DSC_2km_df <- DSC.2km_intersect %>% st_drop_geometry()

z_membership <- function(df, col_name, b_variance) {
  col_name <- enquo(col_name)
  zmf.sum <- c()
  obs.sum <- df %>% pull((!!col_name))
  a = min(obs.sum, na.rm = TRUE)
  b = max(obs.sum, na.rm = TRUE) + b_variance
  for (i in 1:length(obs.sum)) {
    if (is.na(obs.sum[i])) {
      zmf.sum[i] = 1
    } else if (obs.sum[i] <= a) {
      zmf.sum[i] = 1
    } else if (obs.sum[i] <= (a + b) / 2) {
      zmf.sum[i] = 1 - 2 * ((obs.sum[i] - a) / (b - a))^2
    } else if (obs.sum[i] <= b) {
      zmf.sum[i] = 2 * ((obs.sum[i] - b) / (b - a))^2
    } else if (obs.sum[i] >= b) {
      zmf.sum[i] = 0
    }
  }
  return(zmf.sum)
}

# get the Z-membership score for each 2km cell
DSC_2km_df$Score.Z_Membership <- z_membership(DSC_2km_df, 'gridcode', 0.001)

DSC_2km_scored_spatial <- grid_2km %>%
  inner_join(DSC_2km_df, by = "CellID_2km")

grid_5km <- grid_test %>% # Assuming grid_test is your 5km grid here
  mutate(cell_area_5km = st_area(geometry))

# calculate weights
DSC_5km_weighted <- sf::st_intersection(DSC_2km_scored_spatial, grid_5km) %>%
  mutate(
    intersect_area = st_area(geometry),
    # Calculate what percentage of the 5km cell this 2km piece occupies
    area_weight = as.numeric(intersect_area) / as.numeric(cell_area_5km),
    # Multiply the 2km Z-score by the area weight
    weighted_z_score = Score.Z_Membership * area_weight
  ) %>%
  st_drop_geometry() %>%

  # 7. Group by the 5km Cell ID and sum the weighted fractions
  group_by(CellID_5km) %>%
  summarize(
    Score.Z_Membership_5km = sum(weighted_z_score, na.rm = TRUE),
    .groups = "drop"
  )

DSC_RH_z_membership_final <- grid_5km %>%
  inner_join(DSC_5km_weighted, by = "CellID_5km") %>%
  select(-cell_area_5km) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  DSC_RH_z_membership_final,
  "C:\\GitHub\\SMORES\\data\\5km\\DSC_RH_z_membership_scored_full_5km.parquet"
)
```


```{r}
#comparison

new_ALCH <- st_read_parquet(
  "C:\\GitHub\\SMORES\\data\\5km\\ALCH_Ranked_Importance_scored_new.parquet"
) %>%
  filter(Score.ALCH_Ranked_Importance != 0.000)

old_ALCH <- st_read_parquet(
  "C:\\GitHub\\SMORES\\data\\5km\\ALCH_Ranked_Importance_scored_full.parquet"
)

plot_new_ALCH <- ggplot(data = new_ALCH) +
  geom_sf(aes(fill = Score.ALCH_Ranked_Importance))

plot_new_ALCH

plot_old_ALCH <- ggplot(data = old_ALCH) +
  geom_sf(aes(fill = Score.ALCH_Ranked_Importance))

plot_old_ALCH
# visual check looks like we need to remove values of 0

new_ALCO <- st_read_parquet(
  "C:\\GitHub\\SMORES\\data\\5km\\ALCO_Ranked_Importance_scored_new.parquet"
) |>
  filter(Score.ALCO_Ranked_Importance != 0.000)
old_ALCO <- st_read_parquet(
  "C:\\GitHub\\SMORES\\data\\5km\\ALCO_Ranked_Importance_scored_full.parquet"
)

plot_new_ALCO <- ggplot(data = new_ALCO) +
  geom_sf(aes(fill = Score.ALCO_Ranked_Importance))

plot_new_ALCO

plot_old_ALCO <- ggplot(data = old_ALCO) +
  geom_sf(aes(fill = Score.ALCO_Ranked_Importance))

plot_old_ALCO
```