<!-- # Data Processing for SMORES continuous datasets for the 5km grid 

Methodology: Area Weighted Average

Plan: each dataset will have its own code chunk. All processing will live in that code chunk. The top of the document will house functions, data pathways, and other pieces that will be utilised throughout the script. 


```{r setup}
#load libraries
library(tidyverse)
library(sf)
library(arcgis)
library(arrow)
library(sfarrow)

# 2km base grid
grid.2km <- sf::st_read(
  dsn = "G:\\Shared drives\\NMFS NWC FRAM GIS\\GIS_Base\\Grid.gdb",
  layer = "grid2km_ETsquare"
)

# st_write_parquet(
#   grid.2km,
#   "C:\\GitHub\\SMORES\\data\\2km\\grid_full.parquet"
# )

# get base grid as a dataframe
base_grid_df <- grid.5km %>%
  st_drop_geometry()

# write_parquet(
#   base_grid_df,
#   "C:\\GitHub\\SMORES\\data\\5km\\base_grid_df.parquet"
# )

# calculate the area of the grid cells
grid_2km <- grid.2km %>%
  mutate(cell_area = st_area(.))

# set standard coordinate reference system if using 2 km square grid
crsOut <- st_crs(grid.2km)

Fisheries <- sf::st_read(
  "G:\\My Drive\\SMORES\\NCCOSfisherieslayers20250815081621\\Non-Confidential_NMFS_fisheries_submodel_data_2km_grid.shp"
) %>%
  st_transform(crsOut)

Trawl_Fisheries <- sf::st_read(
  "G:\\My Drive\\SMORES\\Scenario_4_trawl_polygon\\Scenario_4_trawl_polygon.shp"
)

#convert each layer to sf object
melissa_file_path <- "G:\\Shared drives\\NMFS NWC OEI GIS\\ArcGIS\\Projects\\OWEC\\p30\\nccos_share_CA2.gdb"
```


# Deep Sea Coral Robust High Suitability Z Membership
```{r DSC}
DSC.RobustHigh <- sf::st_read(
  dsn = melissa_file_path,
  layer = "DSC_HabitatSuitability_RobustHigh"
) %>%
  st_transform(crsOut)

DSC.2km_intersect <- sf::st_intersection(DSC.RobustHigh, grid_2km) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>%
  # keep only the largest overlapping piece, prevent ties from duplicating rows
  slice_max(area.part, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(CellID_2km, gridcode)

# rop geometry to run the Z-membership math efficiently
DSC_2km_df <- DSC.2km_intersect %>% st_drop_geometry()

z_membership <- function(df, col_name, b_variance) {
  col_name <- enquo(col_name)
  zmf.sum <- c()
  obs.sum <- df %>% pull((!!col_name))
  a = min(obs.sum, na.rm = TRUE)
  b = max(obs.sum, na.rm = TRUE) + b_variance
  for (i in 1:length(obs.sum)) {
    if (is.na(obs.sum[i])) {
      zmf.sum[i] = 1
    } else if (obs.sum[i] <= a) {
      zmf.sum[i] = 1
    } else if (obs.sum[i] <= (a + b) / 2) {
      zmf.sum[i] = 1 - 2 * ((obs.sum[i] - a) / (b - a))^2
    } else if (obs.sum[i] <= b) {
      zmf.sum[i] = 2 * ((obs.sum[i] - b) / (b - a))^2
    } else if (obs.sum[i] >= b) {
      zmf.sum[i] = 0
    }
  }
  return(zmf.sum)
}

# get the Z-membership score for each 2km cell
DSC_2km_df$Score.Z_Membership <- z_membership(DSC_2km_df, 'gridcode', 0.001)

DSC_2km_scored_spatial <- grid_2km %>%
  inner_join(DSC_2km_df, by = "CellID_2km")

st_write_parquet(
  DSC_RH_z_membership_final,
  "C:\\GitHub\\SMORES\\data\\2km\\DSC_RH_z_membership_scored_full_2km.parquet"
)
```

# At-Sea hake mid-water trawl = ASH_RI in fisheries dataset
```{r ASH}
ASH <- Fisheries %>%
  select(ASH_RI)
ASH.grid_RI <- sf::st_intersection(ASH, grid.2km) %>%
  mutate(Score.ASH_Ranked_Importance = ASH_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.ASH_Ranked_Importance) %>% #checked and no duplicates present but there are na values present
  filter(!is.na(Score.ASH_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  ASH.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\ASH_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
ASH.grid <- ASH.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
ASH_scored <- grid_test %>%
  full_join(ASH.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.ASH_Ranked_Importance)) %>%
  rename("0" = Score.ASH_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "ASH",
    values_to = "Score.ASH"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-ASH)

st_write_parquet(
  ASH_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\ASH_scored_full.parquet"
)
```

# Shoreside hake mid-water trawl = SSH_RI in fisheries dataset
```{r SSH}
SSH <- Fisheries %>%
  select(SSH_RI)
SSH.grid_RI <- sf::st_intersection(SSH, grid.2km) %>%
  mutate(Score.SSH_Ranked_Importance = SSH_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.SSH_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.SSH_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  SSH.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\SSH_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
SSH.grid <- SSH.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
SSH_scored <- grid_test %>%
  full_join(SSH.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.SSH_Ranked_Importance)) %>%
  rename("0" = Score.SSH_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "SSH",
    values_to = "Score.SSH"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-SSH)

st_write_parquet(
  SSH_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\SSH_scored_full.parquet"
)
```

# Groundfish bottom trawl = GFBT_RI in fisheries dataset
```{r GFBT}
GFBT <- Fisheries %>%
  select(GFBT_RI)
GFBT.grid_RI <- sf::st_intersection(GFBT, grid_test) %>%
  mutate(Score.GFBT_Ranked_Importance = GFBT_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.GFBT_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.GFBT_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  GFBT.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\GFBT_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
GFBT.grid <- GFBT.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
GFBT_scored <- grid_test %>%
  full_join(GFBT.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.GFBT_Ranked_Importance)) %>%
  rename("0" = Score.GFBT_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "GFBT",
    values_to = "Score.GFBT"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-GFBT)

st_write_parquet(
  GFBT_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\GFBT_scored_full.parquet"
)
```

# Groundfish pot gear = GFP_RI in fisheries dataset
```{r GFP}
GFP <- Fisheries %>%
  select(GFP_RI)
GFP.grid_RI <- sf::st_intersection(GFP, grid.2km) %>%
  mutate(Score.GFP_Ranked_Importance = GFP_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.GFP_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.GFP_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  GFP.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\GFP_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
GFP.grid <- GFP.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
GFP_scored <- grid_test %>%
  full_join(GFP.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.GFP_Ranked_Importance)) %>%
  rename("0" = Score.GFP_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "GFP",
    values_to = "Score.GFP"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-GFP)

st_write_parquet(
  GFP_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\GFP_scored_full.parquet"
)
```

# Groundfish longline gear = GFLL_RI in fisheries dataset
```{r GFLL}
GFLL <- Fisheries %>%
  select(GFLL_RI)
GFLL.grid_RI <- sf::st_intersection(GFLL, grid.2km) %>%
  mutate(Score.GFLL_Ranked_Importance = GFLL_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.GFLL_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.GFLL_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  GFLL.grid_RI,
  "C:\\GitHub\\SMORES\\data\\5km\\GFLL_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
GFLL.grid <- GFLL.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
GFLL_scored <- grid_test %>%
  full_join(GFLL.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.GFLL_Ranked_Importance)) %>%
  rename("0" = Score.GFLL_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "GFLL",
    values_to = "Score.GFLL"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-GFLL)

st_write_parquet(
  GFLL_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\GFLL_scored_full.parquet"
)
```

# Pink shrimp trawl = PS_RI in fisheries dataset
```{r PS}
PS <- Fisheries %>%
  select(PS_RI)

PS.grid_RI <- sf::st_intersection(PS, grid.2km) %>%
  mutate(Score.PS_Ranked_Importance = PS_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.PS_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.PS_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')
st_crs(DSC_RH_scored_long) == st_crs(PS.grid_RI)

st_write_parquet(
  PS.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\PS_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
PS.grid <- PS.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
PS_scored <- grid_test %>%
  full_join(PS.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.PS_Ranked_Importance)) %>%
  rename("0" = Score.PS_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "PS",
    values_to = "Score.PS"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-PS)

st_write_parquet(
  PS_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\PS_scored_full.parquet"
)
```

# Dungeness Crab = CRAB_RI in fisheries dataset
```{r CRAB}
CRAB <- Fisheries %>%
  select(CRAB_RI)
CRAB.grid_RI <- sf::st_intersection(CRAB, grid.2km) %>%
  mutate(Score.CRAB_Ranked_Importance = CRAB_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.CRAB_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.CRAB_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  CRAB.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\CRAB_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
CRAB.grid <- CRAB.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
CRAB_scored <- grid_test %>%
  full_join(CRAB.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.CRAB_Ranked_Importance)) %>%
  rename("0" = Score.CRAB_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "CRAB",
    values_to = "Score.CRAB"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-CRAB)

st_write_parquet(
  CRAB_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\CRAB_scored_full.parquet"
)
```

# Commercial troll/hook and line albacore = ALCO_RI in fisheries dataset
```{r ALCO}
ALCO <- Fisheries %>%
  select(ALCO_RI)
ALCO.grid_RI <- sf::st_intersection(ALCO, grid.2km) %>%
  mutate(Score.ALCO_Ranked_Importance = ALCO_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.ALCO_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.ALCO_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  ALCO.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\ALCO_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
ALCO.grid <- ALCO.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
ALCO_scored <- grid_test %>%
  full_join(ALCO.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.ALCO_Ranked_Importance)) %>%
  rename("0" = Score.ALCO_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "ALCO",
    values_to = "Score.ALCO"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-ALCO)

st_write_parquet(
  ALCO_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\ALCO_scored_full.parquet"
)
```

# Charter vessel albacore troll/hook and line = ALCH_RI in fisheries dataset
```{r ALCH}
ALCH <- Fisheries %>%
  select(ALCH_RI)
ALCH.grid_RI <- sf::st_intersection(ALCH, grid_test) %>%
  mutate(Score.ALCH_Ranked_Importance = ALCH_RI) %>%
  mutate(area.part = st_area(.)) %>%
  group_by(CellID_2km) %>% #use for 2km grid
  slice_max(area.part, n = 1) %>%
  select(CellID_2km, Score.ALCH_Ranked_Importance) %>% #checked and no duplicates present
  filter(!is.na(Score.ALCH_Ranked_Importance)) %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

st_write_parquet(
  ALCH.grid_RI,
  "C:\\GitHub\\SMORES\\data\\2km\\ALCH_Ranked_Importance_scored_full.parquet"
)

# to make the user selected score options
ALCH.grid <- ALCH.grid_RI %>%
  st_drop_geometry() %>%
  group_by(CellID_2km) %>%
  distinct(CellID_2km, .keep_all = TRUE)
ALCH_scored <- grid_test %>%
  full_join(ALCH.grid, by = "CellID_2km") %>%
  filter(!is.na(Score.ALCH_Ranked_Importance)) %>%
  rename("0" = Score.ALCH_Ranked_Importance) %>%
  mutate("0" = 0, "0.01" = 0.01, "0.001" = 0.001) %>%
  pivot_longer(
    cols = starts_with(c("0.", 0)),
    names_to = "ALCH",
    values_to = "Score.ALCH"
  ) %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  select(-ALCH)

st_write_parquet(
  ALCH_scored,
  "C:\\GitHub\\SMORES\\data\\2km\\ALCH_scored_full.parquet"
)
```